---
title: "CNB Lesion/Structural/PRL Final Analyses (Post-Replication)"
author: "Erica Baller"
date: "2026-09-17"
output:
  html_document: default
  pdf_document: default
---

<!--
Pre: make_combined_spreadsheet_for_cnb_paper.Rmd -- reads that script's deidentified
     output (cnb_prl_mimosa_struct_vol_deid.csv), which carries only the columns this
     doc needs rather than the full combined spreadsheet from make_combined_spreadsheet.Rmd
Post: renders the simplified cognitive domain summary table (lesion count/volume,
      GM/WM volume, thalamic volume, PRL count/volume)
Uses: pared-down extract of cnb_goa_prl_mimosa_streamline_filtering_deid_20260621.Rmd --
      keeps only the chunks that "Cognitive Domain Summary Table (simplified)" actually
      depends on (data read-in/cleaning, kosha z-scoring, cog_domains_imaging,
      cog_imaging_results (thalamus folded into imaging_predictors), prl_domain_results, the summary
      table build). One deliberate trim within a kept chunk: the source doc's
      cog_domain_summary_table chunk also builds and prints a *full* 12-column table
      (col_names/cog_table_for_viewing) before narrowing to the simplified 7-column
      view -- that full-table render is commented out (not deleted) at the end of the
      cog_domain_summary_table chunk below, since it's not needed to produce the
      simplified table but may be useful again later.
Dependencies: Using Rstudio version Version 2024.12.0+467 (2024.12.0+467); R v 4.4.2
              Supplementary Analyses section also needs rstatix, emmeans, ggsignif
              (called with pkg:: rather than library() so nothing gets masked)
CHANGE LOG
  10/9/2026 -- imaging reprocessing (see the merge script's change log + lab notebook 20261009):
    - head-size covariate is now SAMSEG ICV (samseg_icv, a true ICV) for EVERY volume measure,
      instead of FAST CSF+GM+WM (which only counts CSF inside the brain mask). Set in icv_column below
    - Table 2 (final): thalamic volume (SAMSEG), WM lesion count/volume, PRL count/volume. GM/WM are NOT
      reported (associations differ by segmentation tool); SAMSEG GM/WM + both thalamic measures moved to a
      supplementary table (Table2_supp_samseg_gm_wm_thalamus). SAMSEG for both thalamus and ICV
    - FAST rerun on the HD-BET skull strips (fast_hdbet), so the 2 FAST-workaround sensitivity
      analyses at the end (drop 8 FAST outliers; OpenMAP + TBV) are turned off (eval = FALSE), and
      replaced by a segmentation-method comparison section (FAST vs SAMSEG vs OpenMAP)
    - 3T n is 101 (was 100): 24530's T1 was recovered
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tibble)
library(purrr)
library(broom)
library(knitr)
library(kableExtra)

homedir <- "~/Library/CloudStorage/Box-Box/BBL/ms_cnb_20260821/post_replication/"


# shared helper functions (norming, item-level LMs, summary tables, supp ANOVAs) --
# MS-CNB-only subset of msdep_prospective_functions.R, see header of mscnb_functions.R
source(paste0(homedir, "scripts/mscnb_functions.R"))
```

# ============================================================
# COGNITIVE DOMAIN / IMAGING SETUP
# ============================================================

## Define cog names and domains
```{r names_and_domains}

#### ACC FOR EACH TASK
acc_task_columns <- c("cpf_kosha_z_acc", "er40_kosha_z_acc", "lnb2_kosha_z_acc",
                  "spvrt_kosha_z_acc", "pcet_kosha_z_acc", "pcptnl_kosha_z_acc",
                  "pmat_kosha_z_acc", "cpw_kosha_z_acc", "svoltd_kosha_z_acc",
                  "vsplot15_kosha_z_acc", "medf36a_kosha_z_acc", "sctap_kosha_z_acc")

acc_task_columns_names <- c("Facial Memory", "Emotion Identification", "Working Memory",
                            "Verbal Reasoning", "Abstraction", "Attention", "Nonverbal Reasoning",
                            "Verbal Memory", "Spatial Memory", "Spatial Reasoning", "Emotion Differentiation",
                            "Finger Taps")


##### RT FOR EACH TASK
rt_task_columns <- c("cpf_kosha_z_rt", "er40_kosha_z_rt","lnb2_kosha_z_rt",
                  "spvrt_kosha_z_rt", "pcet_kosha_z_rt", "pcptnl_kosha_z_rt",
                  "pmat_kosha_z_rt", "cpw_kosha_z_rt", "svoltd_kosha_z_rt",
                  "vsplot15_kosha_z_rt", "medf36a_kosha_z_rt", "mpract_kosha_z_rt")

rt_task_columns_names <- c("Facial Memory", "Emotion Identification","Working Memory",
                            "Verbal Reasoning", "Abstraction", "Attention", "Nonverbal Reasoning",
                            "Verbal Memory", "Spatial Memory", "Spatial Reasoning", "Emotion Differentiation",
                            "Sensorimotor Speed")

## acc domain columms
ef_acc_task_columns <-c("pcet_kosha_z_acc", "pcptnl_kosha_z_acc", "lnb2_kosha_z_acc")
cc_acc_task_columns <-c("vsplot15_kosha_z_acc", "spvrt_kosha_z_acc","pmat_kosha_z_acc")
mem_acc_task_columns <-c("cpf_kosha_z_acc", "svoltd_kosha_z_acc", "cpw_kosha_z_acc")
emo_acc_task_columns <- c("medf36a_kosha_z_acc", "er40_kosha_z_acc")


#rt domain columns
ef_rt_task_columns <-c("pcet_kosha_z_rt", "pcptnl_kosha_z_rt", "lnb2_kosha_z_rt")
cc_rt_task_columns <-c("vsplot15_kosha_z_rt", "spvrt_kosha_z_rt", "pmat_kosha_z_rt")
mem_rt_task_columns <-c("cpf_kosha_z_rt", "svoltd_kosha_z_rt", "cpw_kosha_z_rt")
emo_rt_task_columns <- c("medf36a_kosha_z_rt","er40_kosha_z_rt")

## cognitive domain labels
domain_labels <- c(
  ef_acc  = "EF Accuracy",
  cc_acc  = "CC Accuracy",
  mem_acc = "Memory Accuracy",
  emo_acc = "Emotion Accuracy",
  ef_rt   = "EF RT",
  cc_rt   = "CC RT",
  mem_rt  = "Memory RT",
  emo_rt  = "Emotion RT"
)

```

## Outcomes and Predictors
```{r outcomes_and_predictors}
## For analyses
cog_outcomes <- c("ef_acc", "cc_acc", "mem_acc", "emo_acc",
                  "ef_rt",  "cc_rt",  "mem_rt",  "emo_rt")

imaging_predictors <- c(
  "volume_of_mimosa_lesions",   # total lesion volume from mimosa
  "lesion_count",               # number of lesions
  "csf_volume",                 # CSF volume
  "gm_volume",                  # grey matter volume
  "wm_volume",                  # white matter volume
  "gm_by_icv",                  # grey matter volume / intracranial volume
  "wm_by_icv",                  # white matter volume / intracranial volume
  "total_volume",                # total brain volume
  "Thalamus_total_openmap_seg", # thalamic volume (OpenMAP-T1)
  "Thalamus_total_by_icv",      # thalamic volume / intracranial volume
  # SAMSEG (added 10/9/2026) -- definitions in the merge script's SAMSEG chunk
  "samseg_gm_total",            # total GM: cortex + deep GM + cerebellar GM
  "samseg_cortical_gm",         # cerebral cortex only
  "samseg_wm_total",            # cerebral + cerebellar WM + WM-hypointensities
  "samseg_thalamus_total"       # thalamus (OpenMAP thalamus is still the main one)
)
```

## Prep for making the imaging table
```{r prep_making_imaging_table}

# Predictors that get intracranial_volume added as a covariate, regardless of which
# outcome (cognitive, depression, anxiety, ...) they're being run against. Applies
# automatically per-predictor inside run_item_imaging_lm/run_item_imaging_lm_by_group
# and the hand-rolled cog_imaging_results* blocks below -- everything else (lesion
# count/volume, total_volume, the already-ICV-normalized gm_by_icv/wm_by_icv/
# Thalamus_total_by_icv, PRL measures) never gets ICV.
icv_covariate_predictors <- c("gm_volume", "wm_volume", "csf_volume", "Thalamus_total_openmap_seg",
                              "samseg_gm_total", "samseg_cortical_gm", "samseg_wm_total", "samseg_thalamus_total")

# which column is the ICV covariate (10/9/2026: SAMSEG's ICV, a true ICV -- brain plus ALL intracranial CSF.
# Was intracranial_volume = FAST CSF+GM+WM, which only counts CSF inside the brain mask). Used by the
# hand-rolled cog_imaging_results block and passed to every run_item_imaging_lm() call below
icv_column <- "samseg_icv"

```


# ============================================================
# DATA READ IN and Z-SCORE CALCULATION
# ============================================================


```{r read_deid_spreadsheet}
combined_final <- read.csv(paste0(homedir, "data/cnb_prl_mimosa_struct_vol_deid.csv"), sep = ",", header = T)

```

# Apply scan and cnb quality filters
```{r scan_quality_filters}
# PRL columns: no filter (all participants)
# Lesion count, lesion volume, T1 volumetrics: bad_scans_native_space == 0
native_space_cols <- c("volume_of_mimosa_lesions", "lesion_count",
                       "csf_volume", "gm_volume", "wm_volume", "gm_by_icv", "wm_by_icv", "total_volume",
                       "Thalamus_total_openmap_seg", "Thalamus_total_by_icv",
                       "samseg_icv", "samseg_gm_total", "samseg_cortical_gm", "samseg_wm_total", "samseg_thalamus_total")
combined_final <- combined_final %>%
  filter(full_cnb == 1) %>%
  mutate(across(all_of(native_space_cols),
                ~ ifelse(bad_scans_native_space != 0, NA_real_, .)))

#being mindful that there are individuals who didn't have a full set or mimosa failed and so numbers change

cat("Total participants:", nrow(combined_final), "\n")
cat("bad_scans_native_space == 0:", sum(combined_final$bad_scans_native_space == 0, na.rm = TRUE), "\n")
cat("bad_scans_native_space == 0 & bad_warps == 0:",
    sum(combined_final$bad_scans_native_space == 0 & combined_final$bad_warps == 0, na.rm = TRUE), "\n")
cat("PRL participants (prl.count not NA):", sum(!is.na(combined_final$prl.count)), "\n")
```

# NORMS
## Read in norms
```{r read_in_kosha_norms}
#read in norms by age
cnb_norms_from_kosha <- read.csv(paste0(homedir, "/data/cnb_norm_data.csv"), header = T, sep = ",", fileEncoding = "UTF-8-BOM")

```

## Z-score calculations
```{r r_ms_z_scores_kosha}
# Define the task list - COLUMNS CHANGED TO MATCH
ms_tasks_kosha <- list(
  er40 = c("ER40_A.ER40_CR", "ER40_A.ER40_RTCR"),
  cpf = c("CPF_A.CPF_CR", "CPF_A.CPF_RTCR"),
 # cpfd = c("CPFD_A.CPFD_CR", "CPFD_A.CPFD_TPRT"),
  mpract = c("MPRACT.MP2RTCR", "MPRACT.MP2RTCR"), #IGNORE ACCURACY, this task only has RT, so RT in the ACC column to not break algorithm, but it is not used.
  lnb2 = c("LNB2_135.LNB2_PC", "LNB2_135.LNB2_MRTC"),
  spvrt = c("SPVRT_A.SPVRTA_PC", "SPVRT_A.SPVRTA_RTCR"),
  pcet = c("PCET.PCET_ACC2", "PCET.PCET_RTCR"),
  pcptnl = c("PCPTNL.CPT_TPR", "PCPTNL.CPT_TPRT"),
  pmat = c("PMAT24_A.PMAT24_A_PC", "PMAT24_A.PMAT24_A_RTCR"),
  cpw = c("CPW_A.CPW_CR", "CPW_A.CPW_RTCR"),
#  cpwd = c("CPWD_A.CPWD_CR", "CPWD_A.CPWD_W_RTCR"),
  svoltd = c("SVOLT_A.SVOLT_CR", "SVOLT_A.SVOLT_RTCR"),
  vsplot15 = c("VSPLOT15.VSPLOT15_PC", "VSPLOT15.VSPLOT15_RTCR"),
  medf36a = c("MEDF36_A.MEDF36A_PC","MEDF36_A.MEDF36A_RTCR"),
  sctap = c("SCTAP.SCTAP_TOT", "SCTAP.SCTAP_TOT") #IGNORE RT, this task only has ACC, so ACC in the RT column to not break algorithm, but it is not used.

)


for (task in names(ms_tasks_kosha)) {

  cnb_accuracy_col <- ms_tasks_kosha[[task]][1]   # Accuracy column name
  cnb_reaction_time_col <- ms_tasks_kosha[[task]][2]   # Reaction time column name
  print(task)


      # Compute z-scores using norm-based approach
  combined_final[[paste0(task, "_kosha_z_acc")]] <- zscores_kosha_norms(data_frame = combined_final, norm_column_name = cnb_accuracy_col, acc_or_rt = "acc", cnb_column_name = cnb_accuracy_col, norm_df = cnb_norms_from_kosha)

  #multiply RT by -1 so negative scores mean slower time
    combined_final[[paste0(task, "_kosha_z_rt")]] <- (zscores_kosha_norms(data_frame = combined_final, norm_column_name = cnb_reaction_time_col,  acc_or_rt = "rt", cnb_column_name = cnb_reaction_time_col, norm_df = cnb_norms_from_kosha))* -1
}


```

# Compute cog domains on combined_final
```{r cog_domains_with_imaging}
cog_domains_imaging <- combined_final %>%
  mutate(
    mean_acc_z = rowMeans(across(all_of(acc_task_columns)), na.rm = TRUE),
    mean_rt_z  = rowMeans(across(all_of(rt_task_columns)),  na.rm = TRUE),
    ef_acc  = rowMeans(across(all_of(ef_acc_task_columns)),  na.rm = TRUE),
    cc_acc  = rowMeans(across(all_of(cc_acc_task_columns)),  na.rm = TRUE),
    mem_acc = rowMeans(across(all_of(mem_acc_task_columns)), na.rm = TRUE),
    emo_acc = rowMeans(across(all_of(emo_acc_task_columns)), na.rm = TRUE),
    ef_rt   = rowMeans(across(all_of(ef_rt_task_columns)),   na.rm = TRUE),
    cc_rt   = rowMeans(across(all_of(cc_rt_task_columns)),   na.rm = TRUE),
    mem_rt  = rowMeans(across(all_of(mem_rt_task_columns)),  na.rm = TRUE),
    emo_rt  = rowMeans(across(all_of(emo_rt_task_columns)),  na.rm = TRUE)
  )

```


# ============================================================
# COGNITIVE ANALYSES 
# ============================================================
## Total ACC and RT across all tasks

### Compute overall accuracy/RT data

Overall (not per-domain) summary of mean Accuracy vs mean RT, on the same per-subject
`cog_domains_imaging$mean_acc_z`/`$mean_rt_z` used above, via `run_one_sample_t()` so
the bar/CI are self-consistent the same way the domain plots are. This is a two-bar
descriptive summary, not part of the 8-test FDR family above, so no significance
testing/stars here -- just the mean and its 95% CI. Data only in this chunk --
plotting happens below, once the domain data (and the shared y-axis range across all
5 panels) are also available.
```{r overall_acc_rt_data}
overall_acc_rt_data <- bind_rows(
  run_one_sample_t(cog_domains_imaging$mean_acc_z, label = "mean_acc_z") %>% mutate(Metric = "Accuracy"),
  run_one_sample_t(cog_domains_imaging$mean_rt_z,  label = "mean_rt_z")  %>% mutate(Metric = "Response Time")
) %>%
  mutate(
    total_mean_z_score = mean_z,
    total_se = sd_z / sqrt(n),
    ci95_half_width = qt(0.975, df) * total_se
  )

overall_acc_rt_data
```

## Cognitive domain behavioral analyses

### Compute per-domain accuracy/RT one-sample t-tests

One-sample t-test (mean vs 0) per domain x metric, run on the same per-subject domain
z-scores used everywhere else in this doc (`cog_domains_imaging$ef_acc`, `$ef_rt`, etc.
-- the `run_item_imaging_lm` outcomes/`cog_outcomes` from earlier), via the shared
`run_one_sample_t()` helper. All 8 domain x metric tests are shown together on one
figure, so their p-values are FDR-corrected as a single family (same convention as the
imaging tables above) -- `fdr_sig` marks p_fdr < 0.05.

Error bars are the t-distribution 95% CI (`mean +/- qt(0.975, df) * SEM`, using the
same `df` as the one-sample t-test itself, so it's the same CI `t.test()` would report)
rather than raw SEM -- JAMA's figure guidance calls for 95% CI error bars over SEM/SD,
labeled as such on the axis. Note the CI and the FDR star answer different questions:
a CI excluding zero is "significant, uncorrected p<.05"; the star is the FDR-corrected
result across all 8 tests -- keeping both is intentional, not redundant.
```{r domain_acc_rt_data}
domain_metric_outcomes <- list(
  `Executive Function` = list(Accuracy = "ef_acc",  `Response Time` = "ef_rt"),
  `Complex Cognition`     = list(Accuracy = "cc_acc",  `Response Time` = "cc_rt"),
  Memory                  = list(Accuracy = "mem_acc", `Response Time` = "mem_rt"),
  Emotion                 = list(Accuracy = "emo_acc", `Response Time` = "emo_rt")
)

domain_acc_rt_data <- map_dfr(names(domain_metric_outcomes), function(domain) {
  map_dfr(names(domain_metric_outcomes[[domain]]), function(metric) {
    outcome_col <- domain_metric_outcomes[[domain]][[metric]]
    run_one_sample_t(cog_domains_imaging[[outcome_col]], label = outcome_col) %>%
      mutate(Domain = domain, Metric = metric)
  })
}) %>%
  mutate(
    p_fdr   = p.adjust(p, method = "fdr"),
    fdr_sig = ifelse(p_fdr < 0.05, "*", ""),
    total_mean_z_score = mean_z,
    total_se = sd_z / sqrt(n),
    ci95_half_width = qt(0.975, df) * total_se,
    star_y = ifelse(total_mean_z_score >= 0,
                     total_mean_z_score + ci95_half_width + 0.18,
                     total_mean_z_score - ci95_half_width - 0.18)
  )

domain_acc_rt_data
```

### Shared y-axis limits across all 5 plots

Both data frames (`overall_acc_rt_data`, `domain_acc_rt_data`) are now built, so the
shared range for all 5 panels (overall + 4 domains) can be computed once, up front, as
the union of both -- min/max of mean +/- CI, plus the domain plots' `star_y` (room for
the significance asterisks), with a little padding. Using one range for every panel
below keeps them directly comparable instead of each autoscaling on its own.
```{r shared_acc_rt_ylim}
y_pad <- 0.15
shared_acc_rt_ylim <- c(
  min(c(overall_acc_rt_data$total_mean_z_score - overall_acc_rt_data$ci95_half_width,
        domain_acc_rt_data$total_mean_z_score  - domain_acc_rt_data$ci95_half_width,
        domain_acc_rt_data$star_y)) - y_pad,
  max(c(overall_acc_rt_data$total_mean_z_score + overall_acc_rt_data$ci95_half_width,
        domain_acc_rt_data$total_mean_z_score  + domain_acc_rt_data$ci95_half_width,
        domain_acc_rt_data$star_y)) + y_pad
)
```

### Overall accuracy/RT plot

Title dropped (per journal figure convention -- the description belongs in the
manuscript's figure legend, not baked into the image).
```{r total_mean_plot}
total_mean_plot <- ggplot(overall_acc_rt_data, aes(x = Metric, y = total_mean_z_score, fill = Metric)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = total_mean_z_score - ci95_half_width, ymax = total_mean_z_score + ci95_half_width),
                position = position_dodge(width = 0.6), width = 0.2, linewidth = 0.7) +
  labs(y = "Mean z-score (+/- 95% C.I.)", x = "Metric") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14, color = "black"),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    legend.position = "none",
    axis.text = element_text(size = 14, color = "black"),
    axis.line = element_line(size = 1),
    panel.background = element_rect(color = "white"),
    panel.grid = element_blank()
  ) +
  scale_y_continuous(limits = shared_acc_rt_ylim) +
  scale_fill_manual(values = c("Accuracy" = "#5A80C7", "Response Time" = "goldenrod1"))

# Print the plot
total_mean_plot

# Suggested manuscript figure legend:
#   "Overall cognitive performance across all tasks. Bar plot shows the mean z score
#    (+/- 95% confidence interval) for accuracy and response time, averaged across all
#    cognitive tasks, in all participants with CNB data (n = 121). Error bars represent 95%
#    confidence intervals."

# Save the plot
ggsave(paste0(homedir, "results/Figure2.pdf"), plot = total_mean_plot, height = 6, width = 5)
ggsave(paste0(homedir, "results/Figure2.png"), plot = total_mean_plot, height = 6, width = 5)
```

### Build and combine the 4 domain plots
```{r domain_acc_rt_plots}
# plot_data/ylim default to the main-analysis objects; the full-CNB supplement below
# passes its own subsample versions instead
make_domain_acc_rt_plot <- function(domain_label, plot_data = domain_acc_rt_data, ylim = shared_acc_rt_ylim) {
  ggplot(plot_data %>% filter(Domain == domain_label),
         aes(x = Metric, y = total_mean_z_score, fill = Metric)) +
    geom_bar(stat = "identity", width = 0.7) +
    geom_errorbar(aes(ymin = total_mean_z_score - ci95_half_width, ymax = total_mean_z_score + ci95_half_width),
                  position = position_dodge(width = 0.6), width = 0.2, linewidth = 0.7) +
    geom_text(aes(label = fdr_sig, y = star_y), size = 9) +
    labs(y = "Mean z-score (+/- 95% C.I.)", x = "Metric", title = domain_label) +
    theme_classic() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 17, color = "black"),
      axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14, color = "black"),
      axis.text.y = element_text(size = 11, color = "black"),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 14, margin = margin(r = 10)),
      legend.position = "none",
      axis.text = element_text(size = 14, color = "black"),
      axis.line = element_line(size = 1),
      panel.background = element_rect(color = "white"),
      panel.grid = element_blank()
    ) +
    scale_y_continuous(limits = ylim) +
    scale_fill_manual(values = c("Accuracy" = "#5A80C7", "Response Time" = "goldenrod1"))
}

domain_acc_rt_plots <- map(names(domain_metric_outcomes), make_domain_acc_rt_plot)

domain_acc_rt_combined <- wrap_plots(domain_acc_rt_plots, ncol = 2) +
  plot_annotation(
    tag_levels = "A",
    subtitle = "* FDR-corrected (q < 0.05)",
    theme = theme(plot.subtitle = element_text(hjust = 0, size = 12, face = "italic"),
                  plot.tag = element_text(size = 16, face = "bold"))
  )

domain_acc_rt_combined

# Suggested manuscript figure legend:
#   "Cognitive performance by domain, relative to normative expectations. Bar plots
#    show mean age- and sex-normed z scores (+/- 95% confidence interval) for accuracy
#    and response time within each cognitive domain (Executive Function, Complex
#    Cognition, Memory, Emotion) in all participants with CNB data (n = 121). Error bars
#    represent 95% confidence intervals. Asterisks (*) denote domains for which the
#    mean z score differed significantly from 0 (one-sample t test) after false
#    discovery rate (FDR) correction (q < 0.05) across all 8 domain-by-metric
#    comparisons."

# Save the combined figure
ggsave(paste0(homedir, "results/Figure3.pdf"), plot = domain_acc_rt_combined, height = 8, width = 8)
ggsave(paste0(homedir, "results/Figure3.png"), plot = domain_acc_rt_combined, height = 8, width = 8)
```


# ============================================================
# COGNITIVE DOMAIN ANALYSES BY LESION BURDEN
# ============================================================

## Cognitive Domains vs Lesion Burden
```{r cog_by_lesion_burden_dep}
cog_imaging_results <- map(imaging_predictors, function(img_var) {
  # ICV covariate only for GM/WM/CSF volume -- see icv_covariate_predictors above.
  covariates_dep <- c("test_sessions_v.age", "test_sessions_v.gender")
  
  #add icv as a covariate for our prespecified analyses)
  if (img_var %in% icv_covariate_predictors) covariates_dep <- c(covariates_dep, icv_column) #samseg_icv since 10/9/2026

  tibble(outcome = cog_outcomes) %>%
    mutate(
      imaging_measure = img_var,
      model  = map(outcome, ~ lm(reformulate(c(img_var, covariates_dep), response = .x),
                                 data = cog_domains_imaging)),
      img_lm = map(model, ~ tidy(.x, conf.int = TRUE) %>% filter(term == img_var)),
      glance = map(model, glance)
    ) %>%
    transmute(
      imaging_measure,
      outcome,
      estimate  = map_dbl(img_lm, "estimate"),
      std.error = map_dbl(img_lm, "std.error"),
      statistic = map_dbl(img_lm, "statistic"),
      conf.low  = map_dbl(img_lm, "conf.low"),
      conf.high = map_dbl(img_lm, "conf.high"),
      p_value   = map_dbl(img_lm, "p.value"),
      n         = map_dbl(glance, "nobs"),
      df_residual = map_dbl(glance, "df.residual"), # adding df_residual here for Cohen's d calculation in fmt_tp - EAH 9/17/2026
      r2        = map_dbl(glance, "r.squared"),
      adj_r2    = map_dbl(glance, "adj.r.squared"),
      p_fdr     = p.adjust(p_value, method = "fdr"),
      p_bonf    = p.adjust(p_value, method = "bonferroni")
    )
}) %>%
  bind_rows()

cog_imaging_results
```

# ============================================================
# PRL Analyses
# ============================================================

## CNB domains by PRL — full PRL subsample
```{r cnb_vs_prl}
prl_with_cnb <- combined_final %>%
  filter(!is.na(prl.count)) %>%
  mutate(
    mean_acc_z = rowMeans(across(all_of(acc_task_columns)), na.rm = TRUE),
    mean_rt_z  = rowMeans(across(all_of(rt_task_columns)),  na.rm = TRUE),
    ef_acc  = rowMeans(across(all_of(ef_acc_task_columns)),  na.rm = TRUE),
    cc_acc  = rowMeans(across(all_of(cc_acc_task_columns)),  na.rm = TRUE),
    mem_acc = rowMeans(across(all_of(mem_acc_task_columns)), na.rm = TRUE),
    emo_acc = rowMeans(across(all_of(emo_acc_task_columns)), na.rm = TRUE),
    ef_rt   = rowMeans(across(all_of(ef_rt_task_columns)),   na.rm = TRUE),
    cc_rt   = rowMeans(across(all_of(cc_rt_task_columns)),   na.rm = TRUE),
    mem_rt  = rowMeans(across(all_of(mem_rt_task_columns)),  na.rm = TRUE),
    emo_rt  = rowMeans(across(all_of(emo_rt_task_columns)),  na.rm = TRUE)
  )

cat("N in PRL subsample:", nrow(prl_with_cnb), "\n")

cog_domain_outcomes <- c("ef_acc", "cc_acc", "mem_acc", "emo_acc",
                         "ef_rt",  "cc_rt",  "mem_rt",  "emo_rt")

prl_domain_results <- run_item_imaging_lm(cog_domain_outcomes, c("prl.count", "prl.volume..mm3."), prl_with_cnb, icv_covariate_predictors = icv_covariate_predictors, icv_column = icv_column)
prl_domain_results
```

## Cognitive Domain Summary Table

```{r cog_domain_summary_table}

lesion_count_col   <- make_col(cog_imaging_results, "lesion_count")
lesion_volume_col  <- make_col(cog_imaging_results, "volume_of_mimosa_lesions")
csf_volume_col     <- make_col(cog_imaging_results, "csf_volume")
gm_volume_col      <- make_col(cog_imaging_results, "gm_volume")
wm_volume_col      <- make_col(cog_imaging_results, "wm_volume")
gm_by_icv_col      <- make_col(cog_imaging_results, "gm_by_icv")
wm_by_icv_col      <- make_col(cog_imaging_results, "wm_by_icv")
total_volume_col   <- make_col(cog_imaging_results, "total_volume")
prl_count_col      <- make_col(prl_domain_results,  "prl.count")
prl_volume_col     <- make_col(prl_domain_results,  "prl.volume..mm3.")

# Thalamic volume: now folded into imaging_predictors above, so cog_imaging_results
# already carries it (domain-level, its own FDR family across the 8 domains, same as
# lesion/CSF/GM/WM/total/PRL each already get). Thalamus_total_by_icv (thalamic
# volume normalized by ICV) is in imaging_predictors too, same treatment.
thalamus_col            <- make_col(cog_imaging_results, "Thalamus_total_openmap_seg")
thalamus_by_icv_col     <- make_col(cog_imaging_results, "Thalamus_total_by_icv")

# SAMSEG (added 10/9/2026), all adjusted for samseg_icv
samseg_gm_col       <- make_col(cog_imaging_results, "samseg_gm_total")
samseg_ctx_col      <- make_col(cog_imaging_results, "samseg_cortical_gm")
samseg_wm_col       <- make_col(cog_imaging_results, "samseg_wm_total")
samseg_thal_col     <- make_col(cog_imaging_results, "samseg_thalamus_total")

domain_rows <- tibble(outcome = names(domain_labels)) %>%
  mutate(Domain = domain_labels[outcome]) %>%
  left_join(lesion_count_col,  by = "outcome") %>% dplyr::rename(`Lesion Count`    = cell) %>%
  left_join(lesion_volume_col, by = "outcome") %>% dplyr::rename(`Lesion Volume`   = cell) %>%
  left_join(csf_volume_col,    by = "outcome") %>% dplyr::rename(`CSF Volume`      = cell) %>%
  left_join(gm_volume_col,     by = "outcome") %>% dplyr::rename(`GM Volume`       = cell) %>%
  left_join(wm_volume_col,     by = "outcome") %>% dplyr::rename(`WM Volume`       = cell) %>%
  left_join(gm_by_icv_col,     by = "outcome") %>% dplyr::rename(`GM/ICV`          = cell) %>%
  left_join(wm_by_icv_col,     by = "outcome") %>% dplyr::rename(`WM/ICV`          = cell) %>%
  left_join(total_volume_col,  by = "outcome") %>% dplyr::rename(`Total Volume`    = cell) %>%
  left_join(thalamus_col,      by = "outcome") %>% dplyr::rename(`Thalamic Volume` = cell) %>%
  left_join(thalamus_by_icv_col, by = "outcome") %>% dplyr::rename(`Thalamic/ICV`  = cell) %>%
  left_join(samseg_gm_col,     by = "outcome") %>% dplyr::rename(`SAMSEG Total GM`    = cell) %>%
  left_join(samseg_ctx_col,    by = "outcome") %>% dplyr::rename(`SAMSEG Cortical GM` = cell) %>%
  left_join(samseg_wm_col,     by = "outcome") %>% dplyr::rename(`SAMSEG WM`          = cell) %>%
  left_join(samseg_thal_col,   by = "outcome") %>% dplyr::rename(`SAMSEG Thalamus`    = cell) %>%
  left_join(prl_count_col,     by = "outcome") %>% dplyr::rename(`PRL Count`       = cell) %>%
  left_join(prl_volume_col,    by = "outcome") %>% dplyr::rename(`PRL Volume`      = cell) %>%
  select(Domain, `Lesion Count`, `Lesion Volume`,
         `CSF Volume`, `GM Volume`, `WM Volume`, `GM/ICV`, `WM/ICV`, `Total Volume`,
         `Thalamic Volume`, `Thalamic/ICV`,
         `SAMSEG Total GM`, `SAMSEG Cortical GM`, `SAMSEG WM`, `SAMSEG Thalamus`,
         `PRL Count`, `PRL Volume`)

# ---- Overall cognition rows (Mean Accuracy, Mean RT) ----
# Same predictors, but each predictor's FDR family here only ever contains
# these 2 overall outcomes -- run as separate run_item_imaging_lm() calls so
# their p-values never mix into the 8-domain FDR family above. Per Erica: no
# significance star at all on these two rows of mean acc and rt (not part of the domain
# multiple-comparisons family), so a plain t(p), d formatter with no FDR arg.


# NOTE: prl_with_cnb was built fresh in the cnb_vs_prl chunk above (filtered to
# non-NA prl.count, with domain z-scores) -- build our own local copy here too
# rather than depending on that chunk's version still being in the shape we want.
overall_prl_data <- combined_final %>%
  filter(!is.na(prl.count)) %>%
  mutate(
    mean_acc_z = rowMeans(across(all_of(acc_task_columns)), na.rm = TRUE),
    mean_rt_z  = rowMeans(across(all_of(rt_task_columns)),  na.rm = TRUE)
  )

overall_imaging_results  <- run_item_imaging_lm(c("mean_acc_z", "mean_rt_z"), imaging_predictors, cog_domains_imaging, icv_covariate_predictors = icv_covariate_predictors, icv_column = icv_column)
overall_prl_results      <- run_item_imaging_lm(c("mean_acc_z", "mean_rt_z"), c("prl.count", "prl.volume..mm3."), overall_prl_data, icv_covariate_predictors = icv_covariate_predictors, icv_column = icv_column)

overall_labels <- c(mean_acc_z = "Mean Accuracy", mean_rt_z = "Mean RT")

overall_rows <- tibble(outcome = names(overall_labels)) %>%
  mutate(Domain = overall_labels[outcome]) %>%
  left_join(make_col_nostar(overall_imaging_results,  "lesion_count"),               by = "outcome") %>% dplyr::rename(`Lesion Count`    = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "volume_of_mimosa_lesions"),   by = "outcome") %>% dplyr::rename(`Lesion Volume`   = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "csf_volume"),                 by = "outcome") %>% dplyr::rename(`CSF Volume`      = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "gm_volume"),                  by = "outcome") %>% dplyr::rename(`GM Volume`       = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "wm_volume"),                  by = "outcome") %>% dplyr::rename(`WM Volume`       = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "gm_by_icv"),                  by = "outcome") %>% dplyr::rename(`GM/ICV`          = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "wm_by_icv"),                  by = "outcome") %>% dplyr::rename(`WM/ICV`          = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "total_volume"),               by = "outcome") %>% dplyr::rename(`Total Volume`    = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "Thalamus_total_openmap_seg"), by = "outcome") %>% dplyr::rename(`Thalamic Volume` = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "Thalamus_total_by_icv"),      by = "outcome") %>% dplyr::rename(`Thalamic/ICV`    = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "samseg_gm_total"),            by = "outcome") %>% dplyr::rename(`SAMSEG Total GM`    = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "samseg_cortical_gm"),         by = "outcome") %>% dplyr::rename(`SAMSEG Cortical GM` = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "samseg_wm_total"),            by = "outcome") %>% dplyr::rename(`SAMSEG WM`          = cell) %>%
  left_join(make_col_nostar(overall_imaging_results,  "samseg_thalamus_total"),      by = "outcome") %>% dplyr::rename(`SAMSEG Thalamus`    = cell) %>%
  left_join(make_col_nostar(overall_prl_results,      "prl.count"),                  by = "outcome") %>% dplyr::rename(`PRL Count`       = cell) %>%
  left_join(make_col_nostar(overall_prl_results,      "prl.volume..mm3."),           by = "outcome") %>% dplyr::rename(`PRL Volume`      = cell) %>%
  select(Domain, `Lesion Count`, `Lesion Volume`,
         `CSF Volume`, `GM Volume`, `WM Volume`, `GM/ICV`, `WM/ICV`, `Total Volume`,
         `Thalamic Volume`, `Thalamic/ICV`,
         `SAMSEG Total GM`, `SAMSEG Cortical GM`, `SAMSEG WM`, `SAMSEG Thalamus`,
         `PRL Count`, `PRL Volume`)

#overall Mean Accuracy / Mean RT rows first, then the 8 domain rows
cog_summary_table <- bind_rows(overall_rows, domain_rows)

```

## Cognitive Domain Summary Table (simplified)

Table 2 for the paper: thalamic volume (SAMSEG), WM lesion count/volume, and PRL count/volume.

Decision 10/9/2026 (lab notebook 20261009): thalamic volume vs cognition is consistent across segmentation
tools (SAMSEG and OpenMAP give the same FDR-surviving domains), but GM and WM are not -- which tool you use
changes whether GM or WM tracks cognition, most likely because the tools draw the GM/WM boundary differently
(e.g. how T1-dark lesions are classified). GM/WM are also fairly nonspecific, so they are not reported in the
paper; the SAMSEG GM/WM table is kept as a supplement (next chunk) along with the method comparison at the end.
SAMSEG is used for both thalamus and ICV (results match OpenMAP, no advantage to mixing two tools).

```{r cog_domain_summary_table_simple}
col_names_simple <- c(
  "Domain",
  paste0("Thalamic Volume (n=",    get_n(cog_imaging_results, "samseg_thalamus_total"), ")"),
  paste0("WM Lesion Count (n=",    get_n(cog_imaging_results,     "lesion_count"), ")"),
  paste0("WM Lesion Volume (n=",   get_n(cog_imaging_results,     "volume_of_mimosa_lesions"), ")"),
  paste0("PRL Count (n=",       get_n(prl_domain_results,      "prl.count"), ")"),
  paste0("PRL Volume (n=",      get_n(prl_domain_results,      "prl.volume..mm3."), ")")
)

cog_summary_table_simple <- cog_summary_table %>%
  select(Domain, `SAMSEG Thalamus`, `Lesion Count`, `Lesion Volume`, `PRL Count`, `PRL Volume`)

cog_table_for_viewing_simple <- kable(cog_summary_table_simple,
      col.names = col_names_simple,
      caption = "Overall cognition (Mean Accuracy / Mean RT, not part of the domain FDR family, no significance marking) plus cognitive domains (FDR-corrected within each imaging measure, * = FDR p &lt; 0.05) by imaging measures: T(p), d=Cohen's d. Thalamic volume from SAMSEG, adjusted for age, sex, and SAMSEG intracranial volume; lesion and PRL measures adjusted for age and sex",
      align = c("l", rep("c", 5))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
cog_table_for_viewing_simple

# Save the rendered table (with caption/stars/formatting intact) to the results folder.
# self_contained = FALSE skips kableExtra's pandoc-based bundling step (not guaranteed
# to be on PATH in every R setup) -- writes plain HTML plus a small asset subfolder
# instead of a single self-contained file, but needs no external dependency to work.




kableExtra::save_kable(
  cog_table_for_viewing_simple,
  file = paste0(homedir, "results/Table2.png")
)

# Plain-text version of the same table -- same (n=...) headers as the png/kable
# render, so the two stay in sync -- for pulling values into the manuscript
# without re-parsing the image.
cog_summary_table_simple_export <- cog_summary_table_simple
names(cog_summary_table_simple_export) <- col_names_simple
write.csv(cog_summary_table_simple_export,
          file = paste0(homedir, "results/Table2_cog_domain_summary_table_simple.csv"),
          row.names = FALSE)

```

## Supplementary Table: SAMSEG GM/WM and both thalamic measures

Same models as Table 2 (age, sex, + SAMSEG ICV for volumes) with SAMSEG total GM, cortical GM, and total WM,
and thalamic volume from both SAMSEG (Table 2) and OpenMAP-T1. Not in the main paper: GM/WM associations are not
consistent across segmentation tools (see the method comparison section at the end).
```{r supp_table_samseg_gm_wm}
col_names_supp_gm_wm <- c(
  "Domain",
  paste0("Total GM Volume (n=",    get_n(cog_imaging_results, "samseg_gm_total"), ")"),
  paste0("Cortical GM Volume (n=", get_n(cog_imaging_results, "samseg_cortical_gm"), ")"),
  paste0("WM Volume (n=",          get_n(cog_imaging_results, "samseg_wm_total"), ")"),
  paste0("Thalamic Volume, SAMSEG (n=",  get_n(cog_imaging_results, "samseg_thalamus_total"), ")"),
  paste0("Thalamic Volume, OpenMAP (n=", get_n(cog_imaging_results, "Thalamus_total_openmap_seg"), ")")
)

cog_summary_table_supp_gm_wm <- cog_summary_table %>%
  select(Domain, `SAMSEG Total GM`, `SAMSEG Cortical GM`, `SAMSEG WM`, `SAMSEG Thalamus`, `Thalamic Volume`)

cog_table_for_viewing_supp_gm_wm <- kable(cog_summary_table_supp_gm_wm,
      col.names = col_names_supp_gm_wm,
      caption = "SUPPLEMENT: SAMSEG GM/WM volumes and thalamic volume (SAMSEG and OpenMAP-T1), each adjusted for age, sex, and SAMSEG intracranial volume. Overall cognition (Mean Accuracy / Mean RT, not part of the domain FDR family, no significance marking) plus cognitive domains (FDR-corrected within each imaging measure, * = FDR p &lt; 0.05): T(p), d=Cohen's d",
      align = c("l", rep("c", 5))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
cog_table_for_viewing_supp_gm_wm

kableExtra::save_kable(cog_table_for_viewing_supp_gm_wm, file = paste0(homedir, "results/Table2_supp_samseg_gm_wm_thalamus.png"))
cog_summary_table_supp_gm_wm_export <- cog_summary_table_supp_gm_wm
names(cog_summary_table_supp_gm_wm_export) <- col_names_supp_gm_wm
write.csv(cog_summary_table_supp_gm_wm_export,
          file = paste0(homedir, "results/Table2_supp_samseg_gm_wm_thalamus.csv"),
          row.names = FALSE)
```



# ============================================================
# Supplementary Analyses
# ============================================================

## Cognitive domain accuracy anova
```{r supp_acc_domain_anova}
## Runs an anova across all accuracy domains to determine if there are significant differences, and then runs posthoc analyses

# Domains are within-subject (every participant contributes a score to all 4), so
# this is a one-way *repeated-measures* ANOVA (subject as the error stratum), not a
# between-groups ANOVA -- treating the 4 domains as independent groups would ignore
# that the same people are in every group. Sphericity: rstatix::anova_test runs
# Mauchly's test, and get_anova_table(correction = "auto") applies the
# Greenhouse-Geisser correction only when Mauchly's is significant.
# Post hoc: all pairwise domain contrasts from the same RM model via emmeans, with
# Tukey adjustment -- only run when the omnibus ANOVA is significant (p < 0.05).
# Helper functions (make_supp_domain_long, run_supp_domain_anova, make_supp_domain_plot)
# live in mscnb_functions.R; supp_domain_cols below is reused by the RT chunk.

# domain label -> per-subject domain z-score column, in plotting order
supp_domain_cols <- list(
  Accuracy = c(`Executive Function` = "ef_acc", `Complex Cognition` = "cc_acc",
               Memory = "mem_acc", Emotion = "emo_acc"),
  `Response Time` = c(`Executive Function` = "ef_rt", `Complex Cognition` = "cc_rt",
                      Memory = "mem_rt", Emotion = "emo_rt")
)

# ---- Accuracy ----
supp_acc_long  <- make_supp_domain_long(supp_domain_cols$Accuracy, cog_domains_imaging)
supp_acc_anova <- run_supp_domain_anova(supp_acc_long)

cat("Accuracy domains, repeated-measures ANOVA (n =", supp_acc_anova$n, ")\n")
supp_acc_anova$mauchly
supp_acc_anova$anova
if (!is.null(supp_acc_anova$posthoc)) supp_acc_anova$posthoc else cat("Omnibus ANOVA not significant -- no post hoc tests run.\n")

supp_acc_domain_plot <- make_supp_domain_plot(supp_acc_long, supp_acc_anova, fill_color = "#5A80C7")
supp_acc_domain_plot

# Suggested supplementary figure legend:
#   "Accuracy by cognitive domain. Bars show mean age- and sex-normed accuracy z scores
#    (+/- 95% confidence interval) for each cognitive domain in all participants with CNB data (n = 121).
#    Domains were compared with a one-way repeated-measures ANOVA; brackets denote
#    pairwise differences significant after Tukey adjustment
#    (* p < 0.05, ** p < 0.01, *** p < 0.001)."

ggsave(paste0(homedir, "results/supp_fig_acc_domain_anova.pdf"), plot = supp_acc_domain_plot, height = 6, width = 9)
ggsave(paste0(homedir, "results/supp_fig_acc_domain_anova.png"), plot = supp_acc_domain_plot, height = 6, width = 9)

# Plain-text ANOVA + post hoc results for pulling values into the manuscript
write.csv(supp_acc_anova$anova, paste0(homedir, "results/supp_acc_domain_anova.csv"), row.names = FALSE)
if (!is.null(supp_acc_anova$posthoc)) write.csv(supp_acc_anova$posthoc, paste0(homedir, "results/supp_acc_domain_tukey.csv"), row.names = FALSE)
```

## Cognitive domain RT anova
```{r supp_rt_domain_anova}
## Runs an anova across all RT domains to determine if there are significant differences, and then runs posthoc analyses

# Same approach and helpers (mscnb_functions.R) as the accuracy chunk above:
# RM ANOVA, GG correction if sphericity is violated, Tukey post hoc only when the omnibus is significant.
# Reminder: RT z-scores were multiplied by -1 at scoring, so negative = slower.
supp_rt_long  <- make_supp_domain_long(supp_domain_cols$`Response Time`, cog_domains_imaging)
supp_rt_anova <- run_supp_domain_anova(supp_rt_long)

cat("RT domains, repeated-measures ANOVA (n =", supp_rt_anova$n, ")\n")
supp_rt_anova$mauchly
supp_rt_anova$anova
if (!is.null(supp_rt_anova$posthoc)) supp_rt_anova$posthoc else cat("Omnibus ANOVA not significant -- no post hoc tests run.\n")

supp_rt_domain_plot <- make_supp_domain_plot(supp_rt_long, supp_rt_anova, fill_color = "goldenrod1")
supp_rt_domain_plot

# Suggested supplementary figure legend:
#   "Response time by cognitive domain. Bars show mean age- and sex-normed response time
#    z scores (+/- 95% confidence interval; negative values indicate slower responses)
#    for each cognitive domain in all participants with CNB data (n = 121). Domains were compared with a
#    one-way repeated-measures ANOVA; brackets denote pairwise differences significant
#    after Tukey adjustment (* p < 0.05, ** p < 0.01, *** p < 0.001)."

ggsave(paste0(homedir, "results/supp_fig_rt_domain_anova.pdf"), plot = supp_rt_domain_plot, height = 6, width = 9)
ggsave(paste0(homedir, "results/supp_fig_rt_domain_anova.png"), plot = supp_rt_domain_plot, height = 6, width = 9)

write.csv(supp_rt_anova$anova, paste0(homedir, "results/supp_rt_domain_anova.csv"), row.names = FALSE)
if (!is.null(supp_rt_anova$posthoc)) write.csv(supp_rt_anova$posthoc, paste0(homedir, "results/supp_rt_domain_tukey.csv"), row.names = FALSE)
```






# ============================================================
# SENSITIVITY ANALYSIS: DROP 8 SCANS WITH IMPLAUSIBLE FAST VOLUMES
# ============================================================

**TURNED OFF 10/9/2026 (eval = FALSE):** 7 of these 8 had failed BET skull strips. FAST has been rerun on the
HD-BET skull strips (fast_hdbet) and GM/WM/ICV now come from SAMSEG, so this workaround is no longer needed.
Kept for the record; its old outputs (results/Table2_sensitivity_drop8_fast_outliers.*) are from the old FAST run.

FAST gave 8 participants implausible tissue volumes (intracranial volume more than 3.5 robust SDs
from the sample median: 0.25-0.87 million mm3, or 2.3-2.5 million mm3, vs a median of ~1.8 million;
their OpenMAP-T1 thalamic volumes were normal, so the FAST segmentation failed rather than the
brains being unusual). 7 of the 8 are on the commented-out bad_scans_native_space list from the
6/26/2026 QC; 24554 is not on that list (and 24500, who is on the list, has normal FAST volumes, so
is kept here). Checked 10/7/2026. Their lesion values were all within the range of the rest of the
sample.

This re-runs the full Table 2 (domains + overall rows) with those 8 people removed from the 3T
imaging analyses (GM/WM/thalamic volume, lesions). The PRL columns keep everyone (2 of the 8 are in
the 7T PRL sample, but their 7T scans are not affected by the FAST failure), so the PRL columns are
the same as the main Table 2. Same models, covariates (age, sex, + ICV for GM/WM/CSF/thalamus) and FDR families as the
main analysis.
```{r sensitivity_drop_fast_outliers, eval = FALSE}
fast_outlier_bblids <- c(24101, 24313, 24693, 24554, 24238, 24366, 60509, 24305)

cog_domains_imaging_sens <- cog_domains_imaging %>% filter(!bblid %in% fast_outlier_bblids)
# PRL: everyone kept (7T scans unaffected by the FAST failure) -- PRL results below are the main ones
cat("Dropped from the 3T imaging sample:", sum(fast_outlier_bblids %in% cog_domains_imaging$bblid[!is.na(cog_domains_imaging$gm_volume)]),
    " | PRL sample unchanged:", nrow(prl_with_cnb), "\n")

# run_item_imaging_lm gives the same results as the hand-rolled cog_imaging_results block above
# (checked 10/7/2026), so it is used here for the 3T models
cog_imaging_results_sens     <- run_item_imaging_lm(cog_outcomes, imaging_predictors, cog_domains_imaging_sens,
                                                    icv_covariate_predictors = icv_covariate_predictors, icv_column = icv_column)
overall_imaging_results_sens <- run_item_imaging_lm(c("mean_acc_z", "mean_rt_z"), imaging_predictors, cog_domains_imaging_sens,
                                                    icv_covariate_predictors = icv_covariate_predictors, icv_column = icv_column)

# same simplified layout as Table 2: overall rows (no stars, not in the FDR family) then the 8 domains
sens_cols <- list(`GM Volume`       = list(cog_imaging_results_sens, overall_imaging_results_sens, "gm_volume"),
                  `WM Volume`       = list(cog_imaging_results_sens, overall_imaging_results_sens, "wm_volume"),
                  `Thalamic Volume` = list(cog_imaging_results_sens, overall_imaging_results_sens, "Thalamus_total_openmap_seg"),
                  `Lesion Count`    = list(cog_imaging_results_sens, overall_imaging_results_sens, "lesion_count"),
                  `Lesion Volume`   = list(cog_imaging_results_sens, overall_imaging_results_sens, "volume_of_mimosa_lesions"),
                  `PRL Count`       = list(prl_domain_results,       overall_prl_results,          "prl.count"),
                  `PRL Volume`      = list(prl_domain_results,       overall_prl_results,          "prl.volume..mm3."))   # PRL: full sample, same as main

sens_overall_rows <- tibble(outcome = names(overall_labels), Domain = overall_labels)
sens_domain_rows  <- tibble(outcome = names(domain_labels),  Domain = domain_labels)
for (col in names(sens_cols)) {
  sens_overall_rows <- sens_overall_rows %>% left_join(make_col_nostar(sens_cols[[col]][[2]], sens_cols[[col]][[3]]), by = "outcome") %>% dplyr::rename(!!col := cell)
  sens_domain_rows  <- sens_domain_rows  %>% left_join(make_col(sens_cols[[col]][[1]], sens_cols[[col]][[3]]),         by = "outcome") %>% dplyr::rename(!!col := cell)
}
cog_summary_table_sens <- bind_rows(sens_overall_rows, sens_domain_rows) %>% select(-outcome)

col_names_sens <- c("Domain", paste0(c("GM Volume", "WM Volume", "Thalamic Volume", "WM Lesion Count", "WM Lesion Volume", "PRL Count", "PRL Volume"),
                                     " (n=", map_dbl(sens_cols, ~ get_n(.x[[1]], .x[[3]])), ")"))

cog_table_for_viewing_sens <- kable(cog_summary_table_sens,
      col.names = col_names_sens,
      caption = "SENSITIVITY: 8 participants with implausible FAST volumes removed from the 3T imaging columns (PRL columns unchanged, full 7T sample). Overall cognition (Mean Accuracy / Mean RT, not part of the domain FDR family, no significance marking) plus cognitive domains (FDR-corrected within each imaging measure, * = FDR p &lt; 0.05) by imaging measures: T(p), d=Cohen's d",
      align = c("l", rep("c", 7))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
cog_table_for_viewing_sens

kableExtra::save_kable(cog_table_for_viewing_sens, file = paste0(homedir, "results/Table2_sensitivity_drop8_fast_outliers.png"))
cog_summary_table_sens_export <- cog_summary_table_sens
names(cog_summary_table_sens_export) <- col_names_sens
write.csv(cog_summary_table_sens_export,
          file = paste0(homedir, "results/Table2_sensitivity_drop8_fast_outliers.csv"),
          row.names = FALSE)
```



# ============================================================
# SENSITIVITY ANALYSIS: OPENMAP-T1 VOLUMES (EVERYONE KEPT)
# ============================================================

**TURNED OFF 10/9/2026 (eval = FALSE):** this used OpenMAP TBV as a stand-in head-size covariate because there
was no valid ICV. SAMSEG now gives a true ICV, and OpenMAP GM/WM adjusted for it are in the method-comparison
section below. Kept for the record; its old outputs (results/Table2_sensitivity_openmap_volumes_tbv_covariate.*)
are from before the reprocessing.

Second way around the FAST problem: instead of dropping the 8 scans, take GM and WM volume from
OpenMAP-T1 (which segmented all of them sensibly -- see the *_total_openmap_seg columns added in the
merge script, label -> tissue mapping in data/openmap_t1_label_tissue_classes.csv), so every 3T
participant stays in (n=100).
- GM / WM / thalamic volume: OpenMAP-T1, adjusted for age, sex, and TBV.
- TBV (total brain volume) = OpenMAP GM + WM + brainstem (no CSF/ventricles), used ONLY as the size
  covariate in place of ICV (there is no valid ICV -- FAST ICV is broken for the 8). TBV shrinks with
  atrophy, so adjusting for it removes global atrophy: GM/WM/thalamus effects are relative to overall
  brain size (e.g. GM adjusted for TBV ~ GM fraction of the brain), not absolute volume.
- Lesion and PRL columns are unchanged from the main Table 2.
```{r sensitivity_openmap_volumes, eval = FALSE}
cog_domains_imaging_om <- cog_domains_imaging %>%
  mutate(TBV_openmap_seg = GM_total_openmap_seg + WM_total_openmap_seg + Brainstem_total_openmap_seg) %>%
  # same native-space QC rule as the main analysis (bad_scans_native_space is currently empty)
  mutate(across(c(GM_total_openmap_seg, WM_total_openmap_seg, TBV_openmap_seg),
                ~ ifelse(bad_scans_native_space != 0, NA_real_, .)))

om_predictors     <- c("GM_total_openmap_seg", "WM_total_openmap_seg", "Thalamus_total_openmap_seg")
om_icv_predictors <- om_predictors      # all three get the size covariate
om_head_size      <- "TBV_openmap_seg"   # TBV as the size covariate (not a predictor)

cog_imaging_results_om     <- run_item_imaging_lm(cog_outcomes, om_predictors, cog_domains_imaging_om,
                                                  icv_covariate_predictors = om_icv_predictors, icv_column = om_head_size)
overall_imaging_results_om <- run_item_imaging_lm(c("mean_acc_z", "mean_rt_z"), om_predictors, cog_domains_imaging_om,
                                                  icv_covariate_predictors = om_icv_predictors, icv_column = om_head_size)

# same layout as Table 2; lesion/PRL columns come from the main results
om_cols <- list(`GM Volume`       = list(cog_imaging_results_om, overall_imaging_results_om, "GM_total_openmap_seg"),
                `WM Volume`       = list(cog_imaging_results_om, overall_imaging_results_om, "WM_total_openmap_seg"),
                `Thalamic Volume` = list(cog_imaging_results_om, overall_imaging_results_om, "Thalamus_total_openmap_seg"),
                `Lesion Count`    = list(cog_imaging_results,    overall_imaging_results,    "lesion_count"),
                `Lesion Volume`   = list(cog_imaging_results,    overall_imaging_results,    "volume_of_mimosa_lesions"),
                `PRL Count`       = list(prl_domain_results,     overall_prl_results,        "prl.count"),
                `PRL Volume`      = list(prl_domain_results,     overall_prl_results,        "prl.volume..mm3."))

om_overall_rows <- tibble(outcome = names(overall_labels), Domain = overall_labels)
om_domain_rows  <- tibble(outcome = names(domain_labels),  Domain = domain_labels)
for (col in names(om_cols)) {
  om_overall_rows <- om_overall_rows %>% left_join(make_col_nostar(om_cols[[col]][[2]], om_cols[[col]][[3]]), by = "outcome") %>% dplyr::rename(!!col := cell)
  om_domain_rows  <- om_domain_rows  %>% left_join(make_col(om_cols[[col]][[1]], om_cols[[col]][[3]]),         by = "outcome") %>% dplyr::rename(!!col := cell)
}
cog_summary_table_om <- bind_rows(om_overall_rows, om_domain_rows) %>% select(-outcome)

col_names_om <- c("Domain", paste0(c("GM Volume", "WM Volume", "Thalamic Volume", "WM Lesion Count", "WM Lesion Volume", "PRL Count", "PRL Volume"),
                                   " (n=", map_dbl(om_cols, ~ get_n(.x[[1]], .x[[3]])), ")"))

cog_table_for_viewing_om <- kable(cog_summary_table_om,
      col.names = col_names_om,
      caption = "SENSITIVITY: OpenMAP-T1 GM/WM/thalamic volume adjusted for age, sex, and OpenMAP total brain volume (TBV = GM+WM+brainstem) in place of ICV, all 3T participants kept. Lesion and PRL columns as in the main table. Overall cognition (Mean Accuracy / Mean RT, not part of the domain FDR family, no significance marking) plus cognitive domains (FDR-corrected within each imaging measure, * = FDR p &lt; 0.05): T(p), d=Cohen's d",
      align = c("l", rep("c", 7))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
cog_table_for_viewing_om

kableExtra::save_kable(cog_table_for_viewing_om, file = paste0(homedir, "results/Table2_sensitivity_openmap_volumes_tbv_covariate.png"))
cog_summary_table_om_export <- cog_summary_table_om
names(cog_summary_table_om_export) <- col_names_om
write.csv(cog_summary_table_om_export,
          file = paste0(homedir, "results/Table2_sensitivity_openmap_volumes_tbv_covariate.csv"),
          row.names = FALSE)
```



# ============================================================
# SEGMENTATION METHOD COMPARISON: FAST vs SAMSEG vs OpenMAP-T1 (added 10/9/2026)
# ============================================================

Three tools segmented the same 3T T1s (full-CNB sample):
- **FAST** (FSL) -- rerun 10/9/2026 on the HD-BET skull strips, partial-volume-weighted volumes. Not lesion-filled,
  so T1-dark lesions can be called GM.
- **SAMSEG** (FreeSurfer 7.4.1, T1-only) -- main GM/WM/ICV measure. T1 lesions (WM-hypointensities) counted as WM.
- **OpenMAP-T1** -- main thalamus measure; GM/WM totals from its 280 labels (data/openmap_t1_label_tissue_classes.csv).

Part 1 re-runs the Table 2 models with GM/WM/thalamus from each tool, every one adjusted for age, sex, and the
same SAMSEG ICV, so any difference between columns comes from the segmentation, not the covariate.
Part 2 compares the volumes themselves: correlation (do the tools rank people the same way?) and Bland-Altman
(do they agree on absolute size, and does the difference depend on brain size?). Part 3 checks whether FAST GM
runs high in people with more lesions (no lesion filling).
```{r method_comparison_table}
# ---- Part 1: Table 2 models by segmentation method ----
method_predictors <- c(gm_volume = "GM, FAST", GM_total_openmap_seg = "GM, OpenMAP", samseg_gm_total = "GM, SAMSEG",
                       wm_volume = "WM, FAST", WM_total_openmap_seg = "WM, OpenMAP", samseg_wm_total = "WM, SAMSEG",
                       Thalamus_total_openmap_seg = "Thalamus, OpenMAP", samseg_thalamus_total = "Thalamus, SAMSEG")

cog_imaging_results_methods     <- run_item_imaging_lm(cog_outcomes, names(method_predictors), cog_domains_imaging,
                                                       icv_covariate_predictors = names(method_predictors), icv_column = icv_column)
overall_imaging_results_methods <- run_item_imaging_lm(c("mean_acc_z", "mean_rt_z"), names(method_predictors), cog_domains_imaging,
                                                       icv_covariate_predictors = names(method_predictors), icv_column = icv_column)

methods_overall_rows <- tibble(outcome = names(overall_labels), Domain = overall_labels)
methods_domain_rows  <- tibble(outcome = names(domain_labels),  Domain = domain_labels)
for (p in names(method_predictors)) {
  methods_overall_rows <- methods_overall_rows %>% left_join(make_col_nostar(overall_imaging_results_methods, p), by = "outcome") %>% dplyr::rename(!!method_predictors[[p]] := cell)
  methods_domain_rows  <- methods_domain_rows  %>% left_join(make_col(cog_imaging_results_methods, p),             by = "outcome") %>% dplyr::rename(!!method_predictors[[p]] := cell)
}
cog_summary_table_methods <- bind_rows(methods_overall_rows, methods_domain_rows) %>% select(-outcome)

col_names_methods <- c("Domain", paste0(method_predictors, " (n=", map_dbl(names(method_predictors), ~ get_n(cog_imaging_results_methods, .x)), ")"))

cog_table_for_viewing_methods <- kable(cog_summary_table_methods,
      col.names = col_names_methods,
      caption = "SUPPLEMENT: GM, WM, and thalamic volume from each segmentation tool (FAST, OpenMAP-T1, SAMSEG), each adjusted for age, sex, and SAMSEG intracranial volume. Overall cognition (Mean Accuracy / Mean RT, not part of the domain FDR family, no significance marking) plus cognitive domains (FDR-corrected within each imaging measure, * = FDR p &lt; 0.05): T(p), d=Cohen's d",
      align = c("l", rep("c", length(method_predictors)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
cog_table_for_viewing_methods

kableExtra::save_kable(cog_table_for_viewing_methods, file = paste0(homedir, "results/Table2_supp_by_segmentation_method.png"))
cog_summary_table_methods_export <- cog_summary_table_methods
names(cog_summary_table_methods_export) <- col_names_methods
write.csv(cog_summary_table_methods_export,
          file = paste0(homedir, "results/Table2_supp_by_segmentation_method.csv"),
          row.names = FALSE)
```

```{r method_comparison_agreement, fig.width = 12, fig.height = 8}
# ---- Part 2: agreement between the tools' volumes ----
# head size: FAST "ICV" = CSF+GM+WM inside the brain mask; OpenMAP AllLabels = brain + ventricles (shrinks with
# atrophy); SAMSEG ICV = true ICV. These are NOT the same quantity, so expect high r but large mean differences
method_pairs <- tribble(
  ~measure,     ~a_col,                       ~a_name,   ~b_col,                   ~b_name,
  "GM",         "gm_volume",                  "FAST",    "samseg_gm_total",        "SAMSEG",
  "GM",         "GM_total_openmap_seg",       "OpenMAP", "samseg_gm_total",        "SAMSEG",
  "GM",         "gm_volume",                  "FAST",    "GM_total_openmap_seg",   "OpenMAP",
  "WM",         "wm_volume",                  "FAST",    "samseg_wm_total",        "SAMSEG",
  "WM",         "WM_total_openmap_seg",       "OpenMAP", "samseg_wm_total",        "SAMSEG",
  "WM",         "wm_volume",                  "FAST",    "WM_total_openmap_seg",   "OpenMAP",
  "Head size",  "intracranial_volume",        "FAST",    "samseg_icv",             "SAMSEG",
  "Head size",  "AllLabels_total_openmap_seg","OpenMAP", "samseg_icv",             "SAMSEG",
  "Thalamus",   "Thalamus_total_openmap_seg", "OpenMAP", "samseg_thalamus_total",  "SAMSEG"
)

agreement <- pmap_dfr(method_pairs, function(measure, a_col, a_name, b_col, b_name) {
  d <- cog_domains_imaging %>% select(bblid, a = all_of(a_col), b = all_of(b_col)) %>% na.omit()
  diff <- d$b - d$a
  loa  <- mean(diff) + c(-1.96, 1.96) * sd(diff)
  tibble(measure, comparison = paste(b_name, "vs", a_name), n = nrow(d),
         r = cor(d$a, d$b), mean_a = mean(d$a), mean_b = mean(d$b),
         mean_diff = mean(diff), pct_diff = 100 * mean(diff) / mean(d$a),
         loa_low = loa[1], loa_high = loa[2],
         #does the difference grow with size? (slope of difference on the pair mean)
         prop_bias_p = summary(lm(diff ~ I((d$a + d$b) / 2)))$coefficients[2, 4],
         #people outside the limits of agreement -- worth a visual QC
         outside_loa = paste(d$bblid[diff < loa[1] | diff > loa[2]], collapse = " "))
})
agreement %>%
  mutate(across(c(mean_a, mean_b, mean_diff, loa_low, loa_high), round), across(c(r, pct_diff), ~ round(.x, 2)),
         prop_bias_p = signif(prop_bias_p, 2)) %>%
  kable(caption = "Agreement between segmentation tools (full-CNB 3T sample). Volumes in mm3; mean_diff = second tool minus first; LoA = Bland-Altman 95% limits of agreement; prop_bias_p = slope of the difference on the pair mean (proportional bias); outside_loa = bblids outside the limits") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
write.csv(agreement, paste0(homedir, "results/supp_segmentation_method_agreement.csv"), row.names = FALSE)

# Bland-Altman plots: x = mean of the two tools, y = difference (second minus first), lines = mean difference and 95% LoA
ba_plot <- function(measure, a_col, a_name, b_col, b_name) {
  d <- cog_domains_imaging %>% select(a = all_of(a_col), b = all_of(b_col)) %>% na.omit() %>%
    mutate(avg = (a + b) / 2 / 1000, diff = (b - a) / 1000)   # cm3
  m <- mean(d$diff); s <- sd(d$diff)
  ggplot(d, aes(avg, diff)) +
    geom_point(alpha = 0.6) +
    geom_hline(yintercept = m) +
    geom_hline(yintercept = m + c(-1.96, 1.96) * s, linetype = "dashed") +
    labs(title = paste0(measure, ": ", b_name, " vs ", a_name, " (r = ", sprintf("%.2f", cor(d$a, d$b)), ")"),
         x = "Mean of the two (cm3)", y = paste0(b_name, " - ", a_name, " (cm3)")) +
    theme_minimal(base_size = 10)
}
ba_plots <- pmap(method_pairs, ba_plot)
ba_figure <- wrap_plots(ba_plots, ncol = 3)
ba_figure
ggsave(paste0(homedir, "results/supp_segmentation_method_bland_altman.png"), ba_figure, width = 12, height = 10, dpi = 300)
```

```{r method_comparison_lesion_bias}
# ---- Part 3: does FAST GM run high in people with more lesions? ----
# FAST had no lesion filling, so T1-dark lesions can be classified as GM. If so, FAST GM minus SAMSEG GM
# should rise with lesion volume (log, since lesion volume is very skewed). SAMSEG puts T1 lesions in WM.
lesion_bias <- cog_domains_imaging %>%
  filter(!is.na(gm_volume), !is.na(samseg_gm_total), !is.na(volume_of_mimosa_lesions)) %>%
  mutate(fast_minus_samseg_gm = gm_volume - samseg_gm_total,
         log_lesion_volume    = log(volume_of_mimosa_lesions + 1))
summary(lm(fast_minus_samseg_gm ~ log_lesion_volume + samseg_icv, data = lesion_bias))
cor.test(lesion_bias$fast_minus_samseg_gm, lesion_bias$log_lesion_volume)
```
