---
title: "Make df combining cnb/goa/ratings/prl/lesions"
author: "Erica Baller"
date: "2026-06-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
homedir <- "~/Library/CloudStorage/Box-Box/BBL/msdepression/"
resultsdir <- paste0(homedir, "2026_k23_radiology_pull_n104/results/")
```
#Pre: Demographics, CNB, Redcap, ratings, PRL, Lesion volume, Strealine filtering files, mrn->empi, mrn->bblid mapping files
#Post: Deidentified master dataframe that can be used for a variety of analyses. 
#Uses: We will be bringing together a lot of files to make a master spreadsheet that can be used across many analyses. What is really important is that we combine separately so we protect patient data, and then create a deidentified dataset to save and use in other analyses.
#Dependencies: any R will do

## Demos from Oracle
```{r}
demographics <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/select_demos_full_meds.csv"), sep = ",", header = TRUE)
names(demographics)[names(demographics) == "BBLID"] <- "bblid"

#clean
demographics$bblid <- as.character(trimws(demographics$bblid))

```

##CNB
```{r}
# Read in CNB data
cnb_data_v1 <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/cnb_v1_20260202.csv"), sep = ",", header = TRUE)
cnb_data_v2 <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/cnb_v2_20260202.csv"), sep = ",", header = TRUE)
cnb_data_v3 <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/cnb_v3_20260202.csv"), sep = ",", header = TRUE)

#combined_final <- rbind(cnb_data_v1, cnb_data_v2)
cnb_data <- rbind(cnb_data_v1, cnb_data_v2, cnb_data_v3) 

# Remove practice trials and reset the rows
cnb_data <- cnb_data[-c(1, 23, 90, 120), ]
rownames(cnb_data) <- NULL #n=128

names(cnb_data)[names(cnb_data) == "test_sessions.subid"] <- "bblid"
cnb_data$bblid     <- as.character(trimws(cnb_data$bblid))

#clean
cnb_data <- cnb_data %>%
   arrange(bblid, test_sessions_v.dotest) %>%
   distinct(bblid, .keep_all = TRUE)#n=127

names(cnb_data) <- gsub(x=names(cnb_data), pattern = "\\PCET_A\\.", replacement = "PCET\\.")
```

## Redcap (mood/anxiety/physical functioning)
```{r}
## read in redcap data

#there are a couple of people who have missing data - 24092 (skipped some BDI questions); 24253 - missed bdi, 24624 - missed bdi question
redcap_data <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/redcap_data_20260202.csv"), sep = ",", header = TRUE)

#filter ones where scales were not complete
redcap_data <- redcap_data %>%
  filter(scales_validation_complete != 0) %>%
   mutate(bblid     = as.character(bblid), #n=132
         bdi_ii_16 = as.numeric(gsub("[ab]", "", bdi_ii_16)),
         bdi_ii_18 = as.numeric(gsub("[ab]", "", bdi_ii_18)),
         promis_ms_physfunc_15a_calc = rowSums(across(matches("^promis_ms_physfunc_15a_[0-9]+$"))),
         promis_ms_fatigue_8a_calc   = rowSums(across(matches("^promis_ms_fatigue_8a_[0-9]+$")))) %>%
  arrange(bblid, date) %>%
  distinct(bblid, .keep_all = TRUE) #n=127
  
```

## PRL
```{r}
### read in prl data
cnb_and_prl <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/cnb_and_7t_list_2026-mks.csv"), sep = ",", header = TRUE) #n=30
prl_data <- cnb_and_prl %>% filter(!is.na(prl.count)) #n=27
```

## Lesion volumes
```{r}
#read in mimosa volume data
lesion_volume_df <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/mimosa_volume_values_n105_20260702_094134.csv"), header = T, sep = ",")
```

## Lesion count
```{r}
#read in mimosa volume data
lesion_count_df <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/lesion_count_n104.csv"), header = T, sep = ",")
```

## fast segmentation
```{r}
#read in fast data
fast_volumes <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/total_fast_brain_volumes_n104.csv"), header = T, sep = ",")

fast_df <- fast_volumes %>%
  mutate(intracranial_volume = csf_volume + gm_volume + wm_volume,
         gm_by_icv = gm_volume/intracranial_volume,
         wm_by_icv = wm_volume/intracranial_volume)
```

## segmentation volumes
```{r}
#read in openmap-t1 segmentations
#when I first ran the openmap, I had accidentally labeled empi as subject and exam data as session. I ahve changed the script but have not rerun anything. Once this is run from scratch, would just get rid of the rename line

segmentations <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/all_volumes_n105_from_openmap_t1.csv"), header = T, sep = ",") #n=105
segmentations <- rename(segmentations, "EMPI" = "subject", "EXAM_DATE" = "session") ###CHANGE THIS ONCE THIS IS RERUN WITH NEW SCRIPT
segmentations$Thalamus_total <- (segmentations$Thalamus_L + segmentations$Thalamus_R)


names(segmentations) <- ifelse(names(segmentations) %in% c("EMPI", "EXAM_DATE"),
                                names(segmentations),
                                paste0(names(segmentations), "_openmap_seg")) 
```

## Streamline filtering
```{r}
#read in streamline filtering data
fascicle_volumes <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/fascicle_volumes_all_subjects_roi_n104_20260619.csv"), header = T, sep = ",")
```

## Imaging Support files
```{r}
# read in template fascicle volumes to get network sizes
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep = ",", header = T)

# read in depression network grouping (inDepMask == 2 means in depression network)
fascicle_dep_network <- read.csv(paste0(homedir,"templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv"), header = T, sep = ",")

#read in ef network grouping (inEFMask)
fascicle_ef_network <- read.csv(paste0(homedir,"templates/dti/HCP_YA1065_tractography/streamline_volume_within_ef_network_FDR0_01_25_percentile.csv"), header = T, sep = ",")
```

## Conversion files (MRN->EMPI, MRN-> BBLID
```{r}
#read in mrn to bblid
mrn_to_bblid <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/mrn_to_bblid.csv"), sep = ",", header = TRUE)
#a leading byte-order-mark in the csv mangles the first column name (e.g. "X...Participant.Name"); fix the name without changing how the data itself is parsed (fileEncoding = "UTF-8-BOM" was tried but silently truncated the file at a non-UTF8 byte further down, dropping rows)
names(mrn_to_bblid)[1] <- "Participant.Name"

#make lowercase for merging
names(mrn_to_bblid)[names(mrn_to_bblid) == "BBLID"] <- "bblid"

#read mrn to empi
mrn_to_empi <- read.csv(paste0(homedir, "2026_k23_radiology_pull_n104/data/name_mrn_empi.csv"), sep = ",", header = TRUE)
names(mrn_to_empi)[1] <- "NAME"

```

#Combine
##Combine conversion files
```{r}
#remove leading zeros
mrn_to_bblid$MRN_clean <- str_remove(as.character(mrn_to_bblid$MRN), "^0+")
mrn_to_empi$MRN_clean  <- str_remove(as.character(mrn_to_empi$MRN),  "^0+")

mrn_empi_bblid_merged <- merge(mrn_to_bblid, mrn_to_empi, by = "MRN_clean")

#keep NAME (from mrn_to_empi) so it can be dropped at the de-identify step; drop the duplicate Participant.Name from mrn_to_bblid
mrn_empi_bblid <- mrn_empi_bblid_merged %>% select(EMPI, MRN_clean, bblid, NAME)
```

## Combine behavioral data
```{r}
#demo_cnb_data <- cnb_data # keeping this way while I sort out the merging issuemerge(demographics, cnb_data, by = "bblid", all = TRUE) #n=137
demo_cnb_data <- merge(demographics, cnb_data, by = "bblid", all = TRUE) #n=150 - keep all rows
demo_cnb_redcap_data <- merge(demo_cnb_data, redcap_data, by = "bblid", all = TRUE) #n=151 - keep all rows
all_behavioral_data <- merge(demo_cnb_redcap_data, mrn_empi_bblid, by = "bblid") #n=125 - keep only rows that overlap in both groups
names(all_behavioral_data)[names(all_behavioral_data) == "MRN_clean"] <- "MRN"

# drop anyone without actual CNB data -- test_sessions_v.dotest is only populated for a
# completed CNB session, so this catches people who have demographics/imaging but never
# took the CNB battery (e.g. bblid 11117), who the all=TRUE merges above kept around
all_behavioral_data <- all_behavioral_data %>%
  filter(!is.na(test_sessions_v.dotest)) #n=124
```

## Combine PRL data
```{r}
#merge bblid and mrn to prl data, n=27
#prl_with_bblid <- merge(prl_data, mrn_empi_bblid, by.x = "MRN", by.y = "MRN_clean", all.x = TRUE) #%>%
  #select(bblid, prl.count, prl.volume..mm3., COMPLETED.CNB., DATE.COMPLETED.CNB, Closest.7T.scan)

prl_data$MRN <- str_remove(as.character(prl_data$MRN), "^0+")
cnb_redcap_prl_data <- merge(all_behavioral_data, prl_data, by = "MRN", all.x = TRUE)  #n=124

```

## Combine imaging data (mimosa volumes, counts, segmentations, and streamline filtering) with rest of data
```{r}
cnb_redcap_prl_lesion_data <- merge(cnb_redcap_prl_data, lesion_volume_df, by = "EMPI", all.x = TRUE) #n=125

cnb_redcap_prl_lesion_volumes_data <- merge(cnb_redcap_prl_lesion_data, fast_df, by= "EMPI", all.x = TRUE) #n=125

cnb_redcap_prl_lesion_volumes_counts_data <- merge(cnb_redcap_prl_lesion_volumes_data, lesion_count_df, by = "EMPI", all.x = TRUE) #n=125

# add fascicle volumes (proportions: lesioned voxels / total fascicle volume) #of note, 23 people don't have mimosa lesions, while 21 don't have count. However, the count of people with no lesions is 0
cnb_redcap_prl_lesion_volumes_fascicle_volumes_data <- merge(cnb_redcap_prl_lesion_volumes_counts_data, fascicle_volumes, by = "EMPI", all.x = TRUE) #n=125

cnb_redcap_prl_lesion_volumes_segs_fascicle_volumes_data <- merge(cnb_redcap_prl_lesion_volumes_fascicle_volumes_data, segmentations, by = c("EMPI", "EXAM_DATE"), all.x = TRUE) #n=125

#make thalamus by ICV column
cnb_redcap_prl_lesion_volumes_segs_fascicle_volumes_data$Thalamus_total_by_icv <- cnb_redcap_prl_lesion_volumes_segs_fascicle_volumes_data$Thalamus_total_openmap_seg/cnb_redcap_prl_lesion_volumes_segs_fascicle_volumes_data$intracranial_volume

```

#Post processing
## Fascicle data
```{r}
# fascicle names: all columns after EMPI and EXAM_DATE, drop cranial nerves
fascicle_names <- names(fascicle_volumes[3:dim(fascicle_volumes)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

fiber_volume_values_no_cn <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)


```

## Depression/nondepression network
```{r}

#define depression network no cranial nerves
fascicle_dep_network_no_cn <- fascicle_dep_network %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#get names of networks
dep_network_names   <- fascicle_dep_network_no_cn$fascicle[fascicle_dep_network_no_cn$inDepMask == 2]  # top quartile, n=19
nondep_network_names <- fascicle_dep_network_no_cn$fascicle[fascicle_dep_network_no_cn$inDepMask != 2] # bottom 3 quartiles, n=58


#get total volume of networks
dep_network_total_volume    <- sum(fiber_volume_values_no_cn$volume_full[fiber_volume_values_no_cn$fascicle %in% dep_network_names])
nondep_network_total_volume <- sum(fiber_volume_values_no_cn$volume_full[fiber_volume_values_no_cn$fascicle %in% nondep_network_names])
total_volume_all_fascicles  <- dep_network_total_volume + nondep_network_total_volume

# compute per-subject network burden: sum of (proportion * fascicle_volume) / network_total_volume
# fascicle_volumes columns are already proportions (lesioned / total fascicle volume)
# multiply back by fascicle volume to get lesioned voxel count, then divide by network size
dep_fascicle_volumes_lookup    <- fiber_volume_values_no_cn$volume_full[match(dep_network_names,    fiber_volume_values_no_cn$fascicle)]
nondep_fascicle_volumes_lookup <- fiber_volume_values_no_cn$volume_full[match(nondep_network_names, fiber_volume_values_no_cn$fascicle)]

combined_dep <- cnb_redcap_prl_lesion_volumes_segs_fascicle_volumes_data %>%
  rowwise() %>%
  mutate(
    dep_net_prop_injured   = sum(c_across(all_of(dep_network_names))    * dep_fascicle_volumes_lookup,    na.rm = TRUE) / dep_network_total_volume,
    nondep_net_prop_injured = sum(c_across(all_of(nondep_network_names)) * nondep_fascicle_volumes_lookup, na.rm = TRUE) / nondep_network_total_volume,
    total_net_prop_injured  = sum(c_across(all_of(fascicle_names_no_cranial_nerves)) *
                                    fiber_volume_values_no_cn$volume_full[match(fascicle_names_no_cranial_nerves, fiber_volume_values_no_cn$fascicle)],
                                  na.rm = TRUE) / total_volume_all_fascicles
  ) %>%
  ungroup() %>%
  mutate(across(
    c(all_of(fascicle_names_no_cranial_nerves), lesion_count, 
      dep_net_prop_injured, nondep_net_prop_injured, total_net_prop_injured),
    ~ ifelse(is.na(volume_of_mimosa_lesions), NA, .)
  ))

```

## EF/NonEF network
```{r}

#define ef network no cranial nerves
fascicle_ef_network_no_cn <- fascicle_ef_network %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#get names of networks
ef_network_names   <- fascicle_ef_network_no_cn$fascicle[fascicle_ef_network_no_cn$inEFMask == 2]  # top quartile, n=19
nonef_network_names <- fascicle_ef_network_no_cn$fascicle[fascicle_ef_network_no_cn$inEFMask != 2] # bottom 3 quartiles, n=58


#get total volume of networks
ef_network_total_volume    <- sum(fiber_volume_values_no_cn$volume_full[fiber_volume_values_no_cn$fascicle %in% ef_network_names])
nonef_network_total_volume <- sum(fiber_volume_values_no_cn$volume_full[fiber_volume_values_no_cn$fascicle %in% nonef_network_names])
total_volume_all_fascicles  <- ef_network_total_volume + nonef_network_total_volume

# compute per-subject network burden: sum of (proportion * fascicle_volume) / network_total_volume
# fascicle_volumes columns are already proportions (lesioned / total fascicle volume)
# multiply back by fascicle volume to get lesioned voxel count, then divide by network size
ef_fascicle_volumes_lookup    <- fiber_volume_values_no_cn$volume_full[match(ef_network_names,    fiber_volume_values_no_cn$fascicle)]
nonef_fascicle_volumes_lookup <- fiber_volume_values_no_cn$volume_full[match(nonef_network_names, fiber_volume_values_no_cn$fascicle)]

combined_final <- combined_dep %>%
  rowwise() %>%
  mutate(
    ef_net_prop_injured   = sum(c_across(all_of(ef_network_names))    * ef_fascicle_volumes_lookup,    na.rm = TRUE) / ef_network_total_volume,
    nonef_net_prop_injured = sum(c_across(all_of(nonef_network_names)) * nonef_fascicle_volumes_lookup, na.rm = TRUE) / nonef_network_total_volume,
    total_net_prop_injured  = sum(c_across(all_of(fascicle_names_no_cranial_nerves)) *
                                    fiber_volume_values_no_cn$volume_full[match(fascicle_names_no_cranial_nerves, fiber_volume_values_no_cn$fascicle)],
                                  na.rm = TRUE) / total_volume_all_fascicles
  ) %>%
  ungroup() %>%
  mutate(across(
    c(all_of(fascicle_names_no_cranial_nerves),
      ef_net_prop_injured, nonef_net_prop_injured, total_net_prop_injured),
    ~ ifelse(is.na(volume_of_mimosa_lesions), NA, .)
  ))

```



#Cleaning
 - specifically, after QA, some of the scans look bad (skull strip) or the transform looks very bizarre. Will make vectors of those that are unusable at the native space scan, and ones where warp is bad, add to data frame, and save
```{r qa}

#scans were eyeballed 6/26/2026
#bad_scans_native_space <- c("1003024083", "1003553516", "1003731158", "8459090932", "8463547992", "8472631746", "8473414506", "8473706992")
bad_scans_native_space <- c("") 
bad_warps <- c("8000066398","8000230354","100126757","8000037131","8443183371","8444330518","8451646764","8451738538","8466215076","8469129522","8472497932","8477618466","8451053490")

combined_final$bad_scans_native_space <- ifelse(combined_final$EMPI %in% bad_scans_native_space, 1, 0)
combined_final$bad_warps <- ifelse(combined_final$EMPI %in% bad_warps, 1, 0)

```
 
 
    
#De-identify
```{r}
#remove MRN and EMPI, keep BBLID
combined_deidentified <- combined_final %>%
  select(-MRN, -EMPI)

write.csv(combined_deidentified, file = paste0(homedir, "2026_k23_radiology_pull_n104/data/cnb_goa_prl_mimosa_streamline_filtering_dep_net_deid.csv"))
```

#Demographics table
```{r}
# Cleans select_demos_list from above to code for sex and on_dmt variables 
demos_table_data <- combined_deidentified %>%
  # changes 1 & 2 labels to male & female
  mutate(
   
  # changes 0 & 1 binary labels to no & yes for DMT
    on_dmt = factor(on_dmt, levels = c(0,1), labels = c("No", "Yes"))
  )

# separate participants into male and female groups, exclude NA
female_rows <- demos_table_data[demos_table_data$sex == "Female" & !is.na(demos_table_data$sex),]
male_rows <- demos_table_data[demos_table_data$sex == "Male" & !is.na(demos_table_data$sex),]

# N: count how many people are in each sex group and in total where each row = 1 participant 
n_female <- nrow(female_rows)
n_male <- nrow(male_rows)
n_total <- nrow(demos_table_data)

# make first row of table (n count)
n_row <- data.frame(
  Characteristic = "n",
  Female = n_female,
  Male = n_male,
  Total = n_total
)

# Age 
# Calculate average age and standard deviation by gender and for total group
age_row <- data.frame(
  Characteristic = "Age (mean [SD])",
  Female = sprintf("%.2f [%.2f]", mean(female_rows$age, na.rm = TRUE), sd(female_rows$age, na.rm = TRUE)),
  Male = sprintf("%.2f [%.2f]", mean(male_rows$age, na.rm = TRUE), sd(male_rows$age, na.rm = TRUE)),
  Total = sprintf("%.2f [%.2f]", mean(demos_table_data$age, na.rm = TRUE), sd(demos_table_data$age, na.rm = TRUE))
)

# create section header race/ethnicity for table
race_ethnicity_header <- data.frame(
  Characteristic = "Race and Ethnicity", 
  Female = " ", 
  Male = " ", 
  Total = " "
)

# Race & ethnicity breakdown table: cross-tab by subtype, one row per race x ethnicity combo 
# sort race by most common -> least common
race_levels <- names(sort(table(demos_table_data$race),decreasing = TRUE))
# set the ethnicity labels 
ethnicity_levels <- c("Not Hispanic", "Hispanic") 
# create an data frame to hold race x ethnicity levels 
race_ethnicity_subtype <- data.frame()
# for every race x ethnicity combination, count people in each subgroup for female, male, and total 
for (r in race_levels) {
  for (e in ethnicity_levels) {
    # count people in each combination for females, males, and in total 
    n_f <- sum(female_rows$race == r & female_rows$ethnicity == e, na.rm = TRUE)
    n_m <- sum(male_rows$race == r & male_rows$ethnicity == e, na.rm = TRUE)
    n_t <- sum(demos_table_data$race == r & demos_table_data$ethnicity == e, na.rm = TRUE)
    # convert into percentages: divide by total number of each and multiply by 100 
    # percentages are within sex, not within race
    race_ethnicity_subtype <- rbind(race_ethnicity_subtype, data.frame(
      Characteristic = paste0("   ", r, ", ", e),
      Female = paste0(n_f, " (", round(100 * n_f / n_female, 2), "%)"),
      Male   = paste0(n_m, " (", round(100 * n_m / n_male, 2), "%)"),
      Total  = paste0(n_t, " (", round(100 * n_t / n_total, 2), "%)")
    ))
  }
}
 
# Education
# Calculate mean education and standard deviation for each group 
educ_row <- data.frame(
  Characteristic = "Education Level in Years (mean [SD])",
  Female = sprintf("%.2f [%.2f]", mean(female_rows$years_education, na.rm = TRUE), sd(female_rows$years_education, na.rm = TRUE)),
  Male   = sprintf("%.2f [%.2f]", mean(male_rows$years_education, na.rm = TRUE), sd(male_rows$years_education, na.rm = TRUE)),
  Total  = sprintf("%.2f [%.2f]", mean(demos_table_data$years_education, na.rm = TRUE), sd(demos_table_data$years_education, na.rm = TRUE))
)

# On DMT?
# Sum of people from each group on dmt
dmt_female <- sum(female_rows$on_dmt == "Yes", na.rm = TRUE)
dmt_male   <- sum(male_rows$on_dmt == "Yes", na.rm = TRUE)
dmt_total  <- sum(demos_table_data$on_dmt == "Yes", na.rm = TRUE)
 
# create % of people in each group on DMT 
dmt_row <- data.frame(
  Characteristic = "On DMT",
  Female = paste0(dmt_female, " (", round(100 * dmt_female / n_female, 2), "%)"),
  Male   = paste0(dmt_male, " (", round(100 * dmt_male / n_male, 2), "%)"),
  Total  = paste0(dmt_total, " (", round(100 * dmt_total / n_total, 2), "%)")
)
# bind the table rows together and write/save the table
demo_table <- rbind(
  n_row, age_row, race_ethnicity_header, race_ethnicity_subtype, educ_row, dmt_row
)
demo_table
write.csv(demo_table, paste0(resultsdir, "/demographics_table.csv"), row.names = FALSE)
```

## Demographics table by sample (CNB, lesion, structural, PRL)
###Built demo table function
```{r}
# builds one Total-only column (n, age, race/ethnicity, education, on_dmt) for a given subsample
build_demo_col <- function(df, col_name, race_levels, ethnicity_levels) {
  n_total <- nrow(df)

  n_row <- data.frame(Characteristic = "n", Value = as.character(n_total))

  n_female <- sum(df$sex == "Female", na.rm = TRUE)
  female_row <- data.frame(
    Characteristic = "Female",
    Value = paste0(n_female, " (", round(100 * n_female / n_total, 1), "%)")
  )

  age_row <- data.frame(
    Characteristic = "Age (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$age, na.rm = TRUE), sd(df$age, na.rm = TRUE))
  )

  race_ethnicity_header <- data.frame(Characteristic = "Race and Ethnicity", Value = " ")

  race_ethnicity_subtype <- data.frame()
  for (r in race_levels) {
    for (e in ethnicity_levels) {
      n_t <- sum(df$race == r & df$ethnicity == e, na.rm = TRUE)
      race_ethnicity_subtype <- rbind(race_ethnicity_subtype, data.frame(
        Characteristic = paste0("   ", r, ", ", e),
        Value = paste0(n_t, " (", round(100 * n_t / n_total, 1), "%)")
      ))
    }
  }

  educ_row <- data.frame(
    Characteristic = "Education Level in Years (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$years_education, na.rm = TRUE), sd(df$years_education, na.rm = TRUE))
  )

  dmt_total <- sum(df$on_dmt == "Yes", na.rm = TRUE)
  dmt_row <- data.frame(
    Characteristic = "On DMT",
    Value = paste0(dmt_total, " (", round(100 * dmt_total / n_total, 1), "%)")
  )

  physfunc_row <- data.frame(
    Characteristic = "PROMIS Physical Function Total (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$promis_ms_physfunc_15a_calc, na.rm = TRUE), sd(df$promis_ms_physfunc_15a_calc, na.rm = TRUE))
  )

  fatigue_row <- data.frame(
    Characteristic = "PROMIS Fatigue Total (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$promis_ms_fatigue_8a_calc, na.rm = TRUE), sd(df$promis_ms_fatigue_8a_calc, na.rm = TRUE))
  )

  out <- rbind(n_row, female_row, age_row, race_ethnicity_header, race_ethnicity_subtype, educ_row, dmt_row, physfunc_row, fatigue_row)
  names(out)[names(out) == "Value"] <- col_name
  out
}
```

### Supplemental Table 1 
```{r}
# sample definitions
cnb_sample <- demos_table_data %>%
  filter(!is.na(test_sessions_v.dotest))

#lesion_sample <- demos_table_data %>%
 # filter(!is.na(volume_of_mimosa_lesions) & volume_of_mimosa_lesions != 0)

structural_sample <- demos_table_data %>%
  filter(!is.na(gm_volume) & !is.na(wm_volume) & !is.na(Thalamus_total_openmap_seg))

prl_sample <- demos_table_data %>%
  filter(!is.na(prl.count))

cnb_col        <- build_demo_col(cnb_sample, "CNB", race_levels, ethnicity_levels)
#lesion_col     <- build_demo_col(lesion_sample, "Lesion", race_levels, ethnicity_levels)
structural_col <- build_demo_col(structural_sample, "3T", race_levels, ethnicity_levels)
prl_col        <- build_demo_col(prl_sample, "7T", race_levels, ethnicity_levels)

demo_table_by_sample <- cnb_col %>%
  left_join(structural_col, by = "Characteristic") %>%
 # left_join(lesion_col, by = "Characteristic") %>%
  left_join(prl_col, by = "Characteristic")

demo_table_by_sample
write.csv(demo_table_by_sample, paste0(resultsdir, "/demographics_table_by_sample_supp_fig_1.csv"), row.names = FALSE)
```

### Table1
```{r}
# same columns as demo_table_by_sample (CNB, 3T, 7T), trimmed to n, % female, age, education, on DMT
Table1 <- demo_table_by_sample %>%
  filter(Characteristic %in% c("n", "Female", "Age (mean [SD])", "Education Level in Years (mean [SD])", "On DMT"))

Table1
write.csv(Table1, paste0(resultsdir, "/Table1.csv"), row.names = FALSE)
```