---
title: "Make df combining cnb/prl/lesions/structural data (CNB paper subset)"
author: "Erica Baller"
date: "2026-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
homedir <- "~/Library/CloudStorage/Box-Box/BBL/ms_cnb_20260821/post_replication/"
resultsdir <- paste0(homedir, "results/")

```
#Pre: CNB, PRL, Lesion volume, mrn->empi, mrn->bblid mapping files
#Post: Deidentified dataframe containing only the columns needed for
#      cnb_lesion_structural_prl_final_analyses_post_replication_20260917.Rmd (raw CNB
#      task columns for kosha z-scoring, lesion count/volume, GM/WM/CSF/total/ICV
#      volumes, thalamic volume, PRL count/volume, plus bblid/age/gender/QA columns).
#Uses: We will be bringing together a lot of files to make a master spreadsheet that can be used across many analyses. What is really important is that we combine separately so we protect patient data, and then create a deidentified dataset to save and use in other analyses.
#Dependencies: any R will do


## Demos from Oracle
```{r}
demographics <- read.csv(paste0(homedir, "data/select_demos_full_meds.csv"), sep = ",", header = TRUE)
names(demographics)[names(demographics) == "BBLID"] <- "bblid"

#clean
demographics$bblid <- as.character(trimws(demographics$bblid))

```

##CNB
```{r}
# Read in CNB data
cnb_data_v1 <- read.csv(paste0(homedir, "data/cnb_v1_20260202.csv"), sep = ",", header = TRUE)
cnb_data_v2 <- read.csv(paste0(homedir, "data/cnb_v2_20260202.csv"), sep = ",", header = TRUE)
cnb_data_v3 <- read.csv(paste0(homedir, "data/cnb_v3_20260202.csv"), sep = ",", header = TRUE)

#combined_final <- rbind(cnb_data_v1, cnb_data_v2)
cnb_data <- rbind(cnb_data_v1, cnb_data_v2, cnb_data_v3)

# Remove practice trials and reset the rows
cnb_data <- cnb_data[-c(1, 23, 90, 120), ]
rownames(cnb_data) <- NULL #n=128

names(cnb_data)[names(cnb_data) == "test_sessions.subid"] <- "bblid"


#clean -- a few participants are longitudinal (more than one CNB session); keep each person's earliest
#dotest is read in as text (YYYY-MM-DD), so convert to a real date first -- as.Date() errors if the
#format ever changes, rather than silently sorting alphabetically
cnb_data <- cnb_data %>%
   mutate(test_sessions_v.dotest = as.Date(test_sessions_v.dotest)) %>%
   group_by(bblid) %>%
   slice_min(test_sessions_v.dotest, n = 1, with_ties = FALSE) %>%
   ungroup() %>%
   as.data.frame() #n=127

names(cnb_data) <- gsub(x=names(cnb_data), pattern = "\\PCET_A\\.", replacement = "PCET\\.")
```

## Redcap (mood/anxiety/physical functioning)
```{r}
## read in redcap data

#there are a couple of people who have missing data - 24092 (skipped some BDI questions); 24253 - missed bdi, 24624 - missed bdi question
redcap_data <- read.csv(paste0(homedir, "data/redcap_data_20260202.csv"), sep = ",", header = TRUE)

#filter ones where scales were not complete
redcap_data <- redcap_data %>%
  filter(scales_validation_complete != 0) %>%
   mutate(bblid     = as.character(bblid), #n=132
         bdi_ii_16 = as.numeric(gsub("[ab]", "", bdi_ii_16)),
         bdi_ii_18 = as.numeric(gsub("[ab]", "", bdi_ii_18)),
         promis_ms_physfunc_15a_calc = rowSums(across(matches("^promis_ms_physfunc_15a_[0-9]+$"))),
         promis_ms_fatigue_8a_calc   = rowSums(across(matches("^promis_ms_fatigue_8a_[0-9]+$")))) %>%
  #longitudinal participants have more than one redcap row; keep each person's earliest
  #date is read in as text (YYYY-MM-DD) -- convert to a real date first so the ordering is by date, not alphabetical
  mutate(date = as.Date(date),
         msdep_dx_hx_ms_diagnosis = as.Date(na_if(msdep_dx_hx_ms_diagnosis, ""))) %>%
  group_by(bblid) %>%
  #MS diagnosis date (not in the oracle pull) -- the question was only added to redcap in 2025, so for
  #longitudinal participants it can sit on a later visit than the earliest one kept below. Take it from
  #the earliest visit that has one (only 24660 reported two different dates, ~10 days apart)
  arrange(date, .by_group = TRUE) %>%
  mutate(ms_dx_date = first(na.omit(msdep_dx_hx_ms_diagnosis))) %>%
  slice_min(date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  as.data.frame() #n=127
  
```

## PRL
```{r}
### read in prl data
cnb_and_prl <- read.csv(paste0(homedir, "data/cnb_and_7t_list_2026-mks.csv"), sep = ",", header = TRUE) #n=30
prl_data <- cnb_and_prl %>% filter(!is.na(prl.count)) #n=27
```

## Lesion volumes
```{r}
#read in mimosa volume data
lesion_volume_df <- read.csv(paste0(homedir, "data/mimosa_volume_values_n105_20260702_094134.csv"), header = T, sep = ",")
```

## Lesion count
```{r}
#read in mimosa volume data
lesion_count_df <- read.csv(paste0(homedir, "data/lesion_count_n104.csv"), header = T, sep = ",")
```

## fast segmentation
```{r}
#read in fast data
fast_volumes <- read.csv(paste0(homedir, "data/total_fast_brain_volumes_n104.csv"), header = T, sep = ",")

fast_df <- fast_volumes %>%
  mutate(intracranial_volume = csf_volume + gm_volume + wm_volume,
         gm_by_icv = gm_volume/intracranial_volume,
         wm_by_icv = wm_volume/intracranial_volume)
```

## segmentation volumes
```{r}
#read in openmap-t1 segmentations
#when I first ran the openmap, I had accidentally labeled empi as subject and exam data as session. I ahve changed the script but have not rerun anything. Once this is run from scratch, would just get rid of the rename line

segmentations <- read.csv(paste0(homedir, "data/all_volumes_n105_from_openmap_t1.csv"), header = T, sep = ",") #n=105
segmentations <- rename(segmentations, "EMPI" = "subject", "EXAM_DATE" = "session") ###CHANGE THIS ONCE THIS IS RERUN WITH NEW SCRIPT
segmentations$Thalamus_total <- (segmentations$Thalamus_L + segmentations$Thalamus_R)


names(segmentations) <- ifelse(names(segmentations) %in% c("EMPI", "EXAM_DATE"),
                                names(segmentations),
                                paste0(names(segmentations), "_openmap_seg"))
```

## Conversion files (MRN->EMPI, MRN-> BBLID
```{r}
#read in mrn to bblid
mrn_to_bblid <- read.csv(paste0(homedir, "data/mrn_to_bblid.csv"), sep = ",", header = TRUE)
#a leading byte-order-mark in the csv mangles the first column name (e.g. "X...Participant.Name"); fix the name without changing how the data itself is parsed (fileEncoding = "UTF-8-BOM" was tried but silently truncated the file at a non-UTF8 byte further down, dropping rows)
names(mrn_to_bblid)[1] <- "Participant.Name"

#make lowercase for merging
names(mrn_to_bblid)[names(mrn_to_bblid) == "BBLID"] <- "bblid"

#read mrn to empi
mrn_to_empi <- read.csv(paste0(homedir, "data/name_mrn_empi.csv"), sep = ",", header = TRUE)
names(mrn_to_empi)[1] <- "NAME"

#longitudinal participants were pulled once per visit, so a few people are listed more than once as
#identical copies -- without this, the MRN merge below doubles them (60637, 60589, 60587 were each
#in the final spreadsheet twice). Drop exact-duplicate rows only.
mrn_to_empi <- mrn_to_empi[!duplicated(mrn_to_empi), ]

#stop if any MRN is still listed twice, i.e. the same MRN with a different name or EMPI
stopifnot(!any(duplicated(mrn_to_empi$MRN)))

```

#Combine
##Combine conversion files
```{r}
#remove leading zeros
mrn_to_bblid$MRN_clean <- str_remove(as.character(mrn_to_bblid$MRN), "^0+")
mrn_to_empi$MRN_clean  <- str_remove(as.character(mrn_to_empi$MRN),  "^0+")

mrn_empi_bblid_merged <- merge(mrn_to_bblid, mrn_to_empi, by = "MRN_clean")

#drop names (NAME from mrn_to_empi, Participant.Name from mrn_to_bblid) -- only the IDs are needed for merging
mrn_empi_bblid <- mrn_empi_bblid_merged %>% select(EMPI, MRN_clean, bblid)
```

## Combine behavioral data
```{r}
#demo_cnb_data <- cnb_data # keeping this way while I sort out the merging issuemerge(demographics, cnb_data, by = "bblid", all = TRUE) #n=137
demo_cnb_data <- merge(demographics, cnb_data, by = "bblid", all = TRUE) #n=150 - keep all rows
demo_cnb_redcap_data <- merge(demo_cnb_data, redcap_data, by = "bblid", all = TRUE) #n=151 - keep all rows
all_behavioral_data <- merge(demo_cnb_redcap_data, mrn_empi_bblid, by = "bblid") #n=122 - keep only rows that overlap in both groups
names(all_behavioral_data)[names(all_behavioral_data) == "MRN_clean"] <- "MRN"

# drop anyone without actual CNB data -- test_sessions_v.dotest is only populated for a
# completed CNB session, so this catches people who have demographics/imaging but never
# took the CNB battery (e.g. bblid 11117), who the all=TRUE merges above kept around
all_behavioral_data <- all_behavioral_data %>%
  filter(!is.na(test_sessions_v.dotest)) #n=121

# years since MS diagnosis, counted from the date of the study (CNB session) -- NA if no redcap dx date
all_behavioral_data <- all_behavioral_data %>%
  mutate(years_since_ms_dx = as.numeric(test_sessions_v.dotest - ms_dx_date) / 365.25)
```

## Combine PRL data
```{r}
#merge bblid and mrn to prl data, n=27
#prl_with_bblid <- merge(prl_data, mrn_empi_bblid, by.x = "MRN", by.y = "MRN_clean", all.x = TRUE) #%>%
  #select(bblid, prl.count, prl.volume..mm3., COMPLETED.CNB., DATE.COMPLETED.CNB, Closest.7T.scan)

prl_data$MRN <- str_remove(as.character(prl_data$MRN), "^0+")
cnb_redcap_prl_data <- merge(all_behavioral_data, prl_data, by = "MRN", all.x = TRUE)  #n=121

```

## Combine imaging data (mimosa volumes, counts, segmentations) with rest of data
```{r}
cnb_prl_lesion_data <- merge(cnb_redcap_prl_data, lesion_volume_df, by = "EMPI", all.x = TRUE) #n=121

# merge on EMPI + EXAM_DATE (not just EMPI) since fast_df also carries its own EXAM_DATE
# column -- merging on EMPI alone would silently rename both to EXAM_DATE.x/EXAM_DATE.y
# instead of keeping a single column for the later segmentations merge to key on
cnb_prl_lesion_volumes_data <- merge(cnb_prl_lesion_data, fast_df, by = c("EMPI", "EXAM_DATE"), all.x = TRUE) #n=121

cnb_prl_lesion_volumes_counts_data <- merge(cnb_prl_lesion_volumes_data, lesion_count_df, by = "EMPI", all.x = TRUE) #n=121

cnb_prl_lesion_volumes_segs_data <- merge(cnb_prl_lesion_volumes_counts_data, segmentations, by = c("EMPI", "EXAM_DATE"), all.x = TRUE) #n=121

#make thalamus by ICV column
cnb_prl_lesion_volumes_segs_data$Thalamus_total_by_icv <- cnb_prl_lesion_volumes_segs_data$Thalamus_total_openmap_seg/cnb_prl_lesion_volumes_segs_data$intracranial_volume

# 23 people don't have mimosa lesions, while 21 don't have a count -- the count for
# someone with no lesions is 0, not missing, so lesion_count only gets NA'd out when
# there's no lesion volume at all (i.e. the scan/segmentation itself is missing)
combined_final <- cnb_prl_lesion_volumes_segs_data %>%
  mutate(lesion_count = ifelse(is.na(volume_of_mimosa_lesions), NA, lesion_count))

```



#Cleaning
 - specifically, after QA, some of the scans look bad (skull strip) or the transform looks very bizarre. Will make vectors of those that are unusable at the native space scan, and ones where warp is bad, add to data frame, and save
```{r qa}

#scans were eyeballed 6/26/2026
#bad_scans_native_space <- c("1003024083", "1003553516", "1003731158", "8459090932", "8463547992", "8472631746", "8473414506", "8473706992")
bad_scans_native_space <- c("")
bad_warps <- c("8000066398","8000230354","100126757","8000037131","8443183371","8444330518","8451646764","8451738538","8466215076","8469129522","8472497932","8477618466","8451053490")

combined_final$bad_scans_native_space <- ifelse(combined_final$EMPI %in% bad_scans_native_space, 1, 0)
combined_final$bad_warps <- ifelse(combined_final$EMPI %in% bad_warps, 1, 0)

```



#De-identify and select CNB-paper columns
Only the columns that
cnb_lesion_structural_prl_final_analyses_post_replication_20260917.Rmd actually reads:
identifiers/covariates/QA flags, the raw CNB item columns needed for kosha z-scoring,
lesion count/volume, GM/WM/CSF/total/ICV volumes, thalamic volume (raw + ICV-normalized),
and PRL count/volume.
```{r}
cnb_paper_columns <- c(
  # identifiers / covariates / QA
  "bblid", "test_sessions_v.age", "test_sessions_v.gender",
  "bad_scans_native_space", "bad_warps",
  # raw CNB task columns (kosha norm z-scoring, see ms_tasks_kosha in the analysis script)
  "ER40_A.ER40_CR", "ER40_A.ER40_RTCR",
  "CPF_A.CPF_CR", "CPF_A.CPF_RTCR",
  "MPRACT.MP2RTCR",
  "LNB2_135.LNB2_PC", "LNB2_135.LNB2_MRTC",
  "SPVRT_A.SPVRTA_PC", "SPVRT_A.SPVRTA_RTCR",
  "PCET.PCET_ACC2", "PCET.PCET_RTCR",
  "PCPTNL.CPT_TPR", "PCPTNL.CPT_TPRT",
  "PMAT24_A.PMAT24_A_PC", "PMAT24_A.PMAT24_A_RTCR",
  "CPW_A.CPW_CR", "CPW_A.CPW_RTCR",
  "SVOLT_A.SVOLT_CR", "SVOLT_A.SVOLT_RTCR",
  "VSPLOT15.VSPLOT15_PC", "VSPLOT15.VSPLOT15_RTCR",
  "MEDF36_A.MEDF36A_PC", "MEDF36_A.MEDF36A_RTCR",
  "SCTAP.SCTAP_TOT",
  # lesion / structural volumes
  "volume_of_mimosa_lesions", "lesion_count",
  "csf_volume", "gm_volume", "wm_volume", "gm_by_icv", "wm_by_icv",
  "total_volume", "intracranial_volume",
  # thalamic volume
  "Thalamus_total_openmap_seg", "Thalamus_total_by_icv",
  # PRL
  "prl.count", "prl.volume..mm3."
)

cnb_paper_deidentified <- combined_final %>%
  select(all_of(cnb_paper_columns))

write.csv(cnb_paper_deidentified, file = paste0(homedir, "data/cnb_prl_mimosa_struct_vol_deid.csv"))
```

#De-identify
```{r}
#remove MRN, EMPI, and any name columns, keep BBLID
#name columns can come in from more than one source (e.g. the PRL list carries its own, read in
#as "X...NAME" because of a byte-order-mark), so drop by pattern rather than by exact name.
#case-sensitive on purpose -- "NAME"/"Participant.Name" only, so meds_names_only/meds_name_and_dose stay
name_cols <- grep("NAME|Participant\\.Name", names(combined_final), value = TRUE)
combined_deidentified <- combined_final %>%
  select(-MRN, -EMPI, -all_of(name_cols))

#stop before writing if any identifier column survived
stopifnot(!any(grepl("NAME|Participant\\.Name|^MRN|^EMPI", names(combined_deidentified))))

write.csv(combined_deidentified, file = paste0(homedir, "data/cnb_goa_prl_mimosa_streamline_filtering_dep_net_deid.csv"))
```

#Demographics table
```{r}
# Cleans select_demos_list from above to code for sex and on_dmt variables 
demos_table_data <- combined_deidentified %>%
  mutate(
  # changes 0 & 1 binary labels to no & yes for DMT
    on_dmt = factor(on_dmt, levels = c(0,1), labels = c("No", "Yes"))
  )

# separate participants into male and female groups, exclude NA
female_rows <- demos_table_data[demos_table_data$sex == "Female" & !is.na(demos_table_data$sex),]
male_rows <- demos_table_data[demos_table_data$sex == "Male" & !is.na(demos_table_data$sex),]

# N: count how many people are in each sex group and in total where each row = 1 participant 
n_female <- nrow(female_rows)
n_male <- nrow(male_rows)
n_total <- nrow(demos_table_data)

# make first row of table (n count)
n_row <- data.frame(
  Characteristic = "n",
  Female = n_female,
  Male = n_male,
  Total = n_total
)

# Age 
# Calculate average age and standard deviation by gender and for total group
age_row <- data.frame(
  Characteristic = "Age (mean [SD])",
  Female = sprintf("%.2f [%.2f]", mean(female_rows$age, na.rm = TRUE), sd(female_rows$age, na.rm = TRUE)),
  Male = sprintf("%.2f [%.2f]", mean(male_rows$age, na.rm = TRUE), sd(male_rows$age, na.rm = TRUE)),
  Total = sprintf("%.2f [%.2f]", mean(demos_table_data$age, na.rm = TRUE), sd(demos_table_data$age, na.rm = TRUE))
)

# create section header race/ethnicity for table
race_ethnicity_header <- data.frame(
  Characteristic = "Race and Ethnicity", 
  Female = " ", 
  Male = " ", 
  Total = " "
)

# Race & ethnicity breakdown table: cross-tab by subtype, one row per race x ethnicity combo 
# sort race by most common -> least common
race_levels <- names(sort(table(demos_table_data$race),decreasing = TRUE))
# set the ethnicity labels 
ethnicity_levels <- c("Not Hispanic", "Hispanic") 
# create an data frame to hold race x ethnicity levels 
race_ethnicity_subtype <- data.frame()
# for every race x ethnicity combination, count people in each subgroup for female, male, and total 
for (r in race_levels) {
  for (e in ethnicity_levels) {
    # count people in each combination for females, males, and in total 
    n_f <- sum(female_rows$race == r & female_rows$ethnicity == e, na.rm = TRUE)
    n_m <- sum(male_rows$race == r & male_rows$ethnicity == e, na.rm = TRUE)
    n_t <- sum(demos_table_data$race == r & demos_table_data$ethnicity == e, na.rm = TRUE)
    # convert into percentages: divide by total number of each and multiply by 100 
    # percentages are within sex, not within race
    race_ethnicity_subtype <- rbind(race_ethnicity_subtype, data.frame(
      Characteristic = paste0("   ", r, ", ", e),
      Female = paste0(n_f, " (", round(100 * n_f / n_female, 2), "%)"),
      Male   = paste0(n_m, " (", round(100 * n_m / n_male, 2), "%)"),
      Total  = paste0(n_t, " (", round(100 * n_t / n_total, 2), "%)")
    ))
  }
}
 
# Education
# Calculate mean education and standard deviation for each group 
educ_row <- data.frame(
  Characteristic = "Education Level in Years (mean [SD])",
  Female = sprintf("%.2f [%.2f]", mean(female_rows$years_education, na.rm = TRUE), sd(female_rows$years_education, na.rm = TRUE)),
  Male   = sprintf("%.2f [%.2f]", mean(male_rows$years_education, na.rm = TRUE), sd(male_rows$years_education, na.rm = TRUE)),
  Total  = sprintf("%.2f [%.2f]", mean(demos_table_data$years_education, na.rm = TRUE), sd(demos_table_data$years_education, na.rm = TRUE))
)

# Years since MS diagnosis
# mean years from MS diagnosis (redcap) to the study date for each group
dx_row <- data.frame(
  Characteristic = "Years Since MS Diagnosis (mean [SD])",
  Female = sprintf("%.2f [%.2f]", mean(female_rows$years_since_ms_dx, na.rm = TRUE), sd(female_rows$years_since_ms_dx, na.rm = TRUE)),
  Male   = sprintf("%.2f [%.2f]", mean(male_rows$years_since_ms_dx, na.rm = TRUE), sd(male_rows$years_since_ms_dx, na.rm = TRUE)),
  Total  = sprintf("%.2f [%.2f]", mean(demos_table_data$years_since_ms_dx, na.rm = TRUE), sd(demos_table_data$years_since_ms_dx, na.rm = TRUE))
)

# On DMT?
# Sum of people from each group on dmt
dmt_female <- sum(female_rows$on_dmt == "Yes", na.rm = TRUE)
dmt_male   <- sum(male_rows$on_dmt == "Yes", na.rm = TRUE)
dmt_total  <- sum(demos_table_data$on_dmt == "Yes", na.rm = TRUE)
 
# create % of people in each group on DMT 
dmt_row <- data.frame(
  Characteristic = "On DMT",
  Female = paste0(dmt_female, " (", round(100 * dmt_female / n_female, 2), "%)"),
  Male   = paste0(dmt_male, " (", round(100 * dmt_male / n_male, 2), "%)"),
  Total  = paste0(dmt_total, " (", round(100 * dmt_total / n_total, 2), "%)")
)
# bind the table rows together and write/save the table
demo_table <- rbind(
  n_row, age_row, race_ethnicity_header, race_ethnicity_subtype, educ_row, dx_row, dmt_row
)
demo_table
write.csv(demo_table, paste0(resultsdir, "/demographics_table.csv"), row.names = FALSE)
```

## Demographics table by sample (CNB, lesion, structural, PRL)
###Built demo table function
```{r}
# builds one Total-only column (n, age, race/ethnicity, education, on_dmt) for a given subsample
build_demo_col <- function(df, col_name, race_levels, ethnicity_levels) {
  n_total <- nrow(df)

  n_row <- data.frame(Characteristic = "n", Value = as.character(n_total))

  n_female <- sum(df$sex == "Female", na.rm = TRUE)
  female_row <- data.frame(
    Characteristic = "Female",
    Value = paste0(n_female, " (", round(100 * n_female / n_total, 1), "%)")
  )

  age_row <- data.frame(
    Characteristic = "Age (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$age, na.rm = TRUE), sd(df$age, na.rm = TRUE))
  )

  race_ethnicity_header <- data.frame(Characteristic = "Race and Ethnicity", Value = " ")

  race_ethnicity_subtype <- data.frame()
  for (r in race_levels) {
    for (e in ethnicity_levels) {
      n_t <- sum(df$race == r & df$ethnicity == e, na.rm = TRUE)
      race_ethnicity_subtype <- rbind(race_ethnicity_subtype, data.frame(
        Characteristic = paste0("   ", r, ", ", e),
        Value = paste0(n_t, " (", round(100 * n_t / n_total, 1), "%)")
      ))
    }
  }

  educ_row <- data.frame(
    Characteristic = "Education Level in Years (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$years_education, na.rm = TRUE), sd(df$years_education, na.rm = TRUE))
  )

  dx_row <- data.frame(
    Characteristic = "Years Since MS Diagnosis (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$years_since_ms_dx, na.rm = TRUE), sd(df$years_since_ms_dx, na.rm = TRUE))
  )

  dmt_total <- sum(df$on_dmt == "Yes", na.rm = TRUE)
  dmt_row <- data.frame(
    Characteristic = "On DMT",
    Value = paste0(dmt_total, " (", round(100 * dmt_total / n_total, 1), "%)")
  )

  physfunc_row <- data.frame(
    Characteristic = "PROMIS Physical Function Total (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$promis_ms_physfunc_15a_calc, na.rm = TRUE), sd(df$promis_ms_physfunc_15a_calc, na.rm = TRUE))
  )

  fatigue_row <- data.frame(
    Characteristic = "PROMIS Fatigue Total (mean [SD])",
    Value = sprintf("%.1f [%.1f]", mean(df$promis_ms_fatigue_8a_calc, na.rm = TRUE), sd(df$promis_ms_fatigue_8a_calc, na.rm = TRUE))
  )

  out <- rbind(n_row, female_row, age_row, race_ethnicity_header, race_ethnicity_subtype, educ_row, dx_row, dmt_row, physfunc_row, fatigue_row)
  names(out)[names(out) == "Value"] <- col_name
  out
}
```

### Supplemental Table 1 
```{r}
# sample definitions
cnb_sample <- demos_table_data %>%
  filter(!is.na(test_sessions_v.dotest))

#lesion_sample <- demos_table_data %>%
 # filter(!is.na(volume_of_mimosa_lesions) & volume_of_mimosa_lesions != 0)

structural_sample <- demos_table_data %>%
  filter(!is.na(gm_volume) & !is.na(wm_volume) & !is.na(Thalamus_total_openmap_seg))

prl_sample <- demos_table_data %>%
  filter(!is.na(prl.count))

cnb_col        <- build_demo_col(cnb_sample, "CNB", race_levels, ethnicity_levels)
#lesion_col     <- build_demo_col(lesion_sample, "Lesion", race_levels, ethnicity_levels)
structural_col <- build_demo_col(structural_sample, "3T", race_levels, ethnicity_levels)
prl_col        <- build_demo_col(prl_sample, "7T", race_levels, ethnicity_levels)

demo_table_by_sample <- cnb_col %>%
  left_join(structural_col, by = "Characteristic") %>%
 # left_join(lesion_col, by = "Characteristic") %>%
  left_join(prl_col, by = "Characteristic")

demo_table_by_sample
write.csv(demo_table_by_sample, paste0(resultsdir, "/demographics_table_by_sample_supp_fig_1.csv"), row.names = FALSE)
```

### Table1
```{r}
# same columns as demo_table_by_sample (CNB, 3T, 7T), trimmed to n, % female, age, education, years since MS dx, on DMT
Table1 <- demo_table_by_sample %>%
  filter(Characteristic %in% c("n", "Female", "Age (mean [SD])", "Education Level in Years (mean [SD])", "Years Since MS Diagnosis (mean [SD])", "On DMT"))

Table1
write.csv(Table1, paste0(resultsdir, "/Table1.csv"), row.names = FALSE)
```